---
title: "House Prices"
author: "AG"
format: html
---

# Libraries

```{r}
library(themis)
library(probably)
library(vip)
library(skimr)
library(bestNormalize) # for ord QQ norm
library(embed) # for UMAP
library(tidyverse)
```

# First look at the data 

```{r}
data_train <- read_csv("data/train.csv") %>% 
  mutate(across(is_character, ~ as_factor(.x)))

data_test <- read_csv("data/test.csv") %>% 
  mutate(across(is_character, ~ as_factor(.x)))

dim(data_train)
```

I should submit price predictions for the test data set

## Factors

```{r}
data_train %>% 
  skim() %>% 
  yank("factor")
```

Notes on NAs:

- According to `data_description.txt`, NA in `Alley` means "no alley access" - should be replaced by "no access"

- NA in `MasVnrType` means "no data"

- NA in `BsmtQual`means "no basement" - should be replaced

- NA in `BsmtCond` means "no basement" - should be replaced

- NA in `BsmtExposure` means "no basement" - should be replaced

- NA in `BsmtFinType1` means "no basement" - should be replaced

- NA in `BsmtFinType2`means "no basement" - should be replaced

- NA in `Electrical` means "no data"

- NA in `FireplaceQu` means "no fireplace", should be replaced





## Numerics

```{r}
data_train %>% 
  skim() %>% 
  yank("numeric")
```

