---
title: "Stroke prediction"
author: "AG"
date: "last update: `r format(Sys.Date(), format = '%d %B %Y')`"
output:
  html_document:
    toc: yes
    df_print: paged
    code_folding: hide
    fig_width: 10
    fig_height: 6
    theme: cerulean
    highlight: kate
    toc_float: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F, cache = T)
library(tidyverse)
library(caret)
#library(GGally)
library(DMwR) # installed via archive: install.packages("/Path/to/DMwR_0.4.1.tar.gz", repos=NULL, type="source" )
library(ROCR)
```

# The data set

## Read data

In `smoking_status` Unknown should be changed to NA.

Also, it can be ordered: never < formerly < smokes

Other predictors seem to be OK

```{r read}
df <- read_csv("data/healthcare-dataset-stroke-data.csv", col_types = "cfdfffffddcf", na = c("Unknown", "N/A"))
# if you set smoking_status to factor in col_types, na() won't work
df$smoking_status <- as_factor(df$smoking_status)
df$smoking_status <- fct_relevel(df$smoking_status, c("never smoked", "formerly smoked", "smokes"))
df$stroke <- factor(ifelse(df$stroke == 1, "yes", "no"), levels = c("no", "yes"))

df
```

## skimr description

Skip id column

```{r desc}
df$id <- NULL
skimr::skim_to_wide(df)
```

Target 'stroke' is imbalanced!

Smoking's complete rate 0.7

BMI's complete rate 0.96

One 'Other' gender to be removed

```{r}
df <- df %>% filter(gender != "Other")
```


## EDA

### Pairs plot

```{r pairs, fig.width=10, fig.height=10, warning=FALSE, message=FALSE}
GGally::ggpairs(df, aes(color = stroke, alpha = 0.2, dotsize = 0.02), 
        upper = list(continuous = GGally::wrap("cor", size = 2.5)),
        diag = list(continuous = "barDiag")) +
  scale_color_brewer(palette = "Set1", direction = -1) +
  scale_fill_brewer(palette = "Set1", direction = -1)
```

### Detailed

#### Stroke vs Gender

```{r}
gender <- df %>% group_by(stroke, gender) %>% summarize(N=n())

ggplot(gender, aes(stroke, N)) +
  geom_bar(aes(fill=gender), alpha = 0.8, stat = "identity", position = "fill")+
  scale_fill_brewer(palette = "Set2", direction = -1)+
  ylab("proportion")
```

Proportions in both stroke groups are roughly the same

#### Stroke vs Age

```{r}
ggplot(df, aes(stroke, age)) + 
  geom_violin(alpha=0.3) +
  geom_jitter(alpha=0.2, size=0.8, width = 0.15, height = 0.1, aes(color = gender)) + 
  geom_boxplot(alpha = 0.2) +
  scale_color_brewer(palette = "Set2", direction = -1)
```
On average, age is higher among stroke-yes group, there is no interaction between age and gender

#### Stroke vs Hypertension

```{r}
hyptens <- df %>% group_by(stroke, hypertension) %>% summarize(N=n())

hyptens

ggplot(hyptens, aes(stroke, N)) +
  geom_bar(aes(fill=hypertension), alpha = 0.8, stat = "identity", position = "fill")+
  scale_fill_brewer(palette = "Set2", direction = -1)+
  ylab("proportion")
```
Hypertension occurres more often among stroke-yes

## Scaling & Normalization

Scale numeric features

```{r}
df_scaled <- df %>% select(avg_glucose_level, age, bmi) %>% scale() %>% data.frame()
```

## Make Dummies

I've decided to omit smoking_status completely - it won't be *dummified*

```{r}
# select vars
to_dum <- df %>% select(gender, ever_married, work_type, Residence_type)
# make an obj
dummies <- dummyVars(~ ., data=to_dum)
# apply it
df_dummy <- data.frame(predict(dummies, newdata=to_dum))

head(df_dummy)
```

## Join scaled and dummies and the rest

```{r}
df_proc <- bind_cols(df_scaled, df_dummy, select(df, hypertension, heart_disease, stroke))
df_proc
```

## Imputation

I will use median imputation for BMI.

Smoking status will not be processed.

```{r}
df_proc <- df_proc %>% 
  mutate(bmi = ifelse(is.na(bmi), median(bmi, na.rm = TRUE), bmi))
```

# Modelling

## Params tuning

ROC-optimization is better when data is imbalanced

```{r}
fit_ctrl_xgb_roc <- trainControl(## 10-fold CV
                           method = "repeatedcv",
                           number = 5,
                           repeats = 10, 
                           allowParallel = T,
                           classProbs = T,
                           summaryFunction = twoClassSummary)

fit_ctrl_xgb_kappa <- trainControl(## 10-fold CV
                           method = "repeatedcv",
                           number = 5,
                           repeats = 10, 
                           allowParallel = T)
```

## Split data

Imbalanced data - use SMOTE to create training data set, but not testing data set

```{r}
set.seed(1234)
sample_set <- createDataPartition(y = df_proc$stroke, p = .75, list = FALSE)
df_train <- df_proc[sample_set,]
df_test <- df_proc[-sample_set,]

# DMwR::SMOTE for imbalanced data: over=225 and under=150 give me 1:1 ratio
df_train_smote <- SMOTE(stroke ~ ., data.frame(df_train), perc.over = 225, perc.under = 150)

df_train_smote %>% group_by(stroke) %>% summarise(N=n())
```


# RF

## Training and validation

### ROC-optimized

```{r}
set.seed(122)

fit_rf_roc <- train(stroke ~ ., 
                 data = df_train, 
                 metric = "ROC", 
                 method = "rf", 
                 trControl = fit_ctrl_xgb_roc,
                 tuneLength = 10,
                 verbosity = 0,
                 verbose = FALSE)

fit_rf_roc
```

### Kappa-optimized

```{r}
set.seed(123)

fit_rf_acc <- train(stroke ~ ., 
                 data = df_train, 
                 metric = "Kappa", 
                 method = "rf", 
                 trControl = fit_ctrl_xgb_kappa,
                 verbosity = 0,
                 verbose = FALSE)

fit_rf_acc
```


## Features importance

```{r}
imp_vars_rf <- varImp(fit_rf_roc)

plot(imp_vars_rf, main="Variable Importance with RF")

```


## Testing

ROC-optimized model

```{r}
predClasses_rf_roc <- predict(fit_rf_roc, newdata=df_test)

cm_rf_roc <- confusionMatrix(data = predClasses_rf_roc, 
                reference = df_test$stroke,
                mode="everything",
                positive="yes")

cm_rf_roc
```


# XGB

## Training and validation

### ROC optimized

```{r, eval=T}
set.seed(123)

fit_xgb_roc <- train(stroke ~ ., 
                 data = df_train, 
                 method = "xgbDART",
                 metric = "ROC", 
                 trControl = fit_ctrl_xgb_roc,
                 verbose = FALSE,
                 verbosity = 0)

fit_xgb_roc$bestTune
```

Show metrics

```{r, eval=T}
fit_xgb_roc$results %>% filter(nrounds == 50, max_depth == 2, eta == 0.3, gamma == 0, subsample == 1, rate_drop==0.5, colsample_bytree == 0.6, skip_drop == 0.95)
```

## Features importance

```{r, eval=T}
imp_vars_xgb <- varImp(fit_xgb_roc)

plot(imp_vars_xgb, main="Variable Importance with XGB")
```


## Testing

```{r, warning=FALSE}
predClasses_xgb_roc <- predict(fit_xgb_roc, newdata=df_test)

cm_xgb_roc <- confusionMatrix(data = predClasses_xgb_roc, 
                reference = df_test$stroke,
                mode="everything",
                positive="yes")

cm_xgb_roc
```

# ROC & AUC

```{r}
get_roc <- function(fit.obj, testing.df){
  pred_prob <- predict.train(fit.obj, newdata = testing.df, type="prob")
  pred_roc <- prediction(predictions = pred_prob$yes, labels = testing.df$stroke)
  perf_roc <- performance(pred_roc, measure="tpr", x.measure = "fpr")
  return(list(perf_roc, pred_roc))
}
```

## RF 

```{r}
# calculate ROC
perf_pred <- get_roc(fit_rf_roc, df_test)
perf_roc_rf <- perf_pred[[1]]
pred_roc_rf <- perf_pred[[2]]

# take AUC 
auc_rf <- round(unlist(slot(performance(pred_roc_rf, measure = "auc"), "y.values")), 3)

# plot
plot(perf_roc_rf, main = "RF ROC curve", col = "steelblue", lwd = 3)
abline(a = 0, b = 1, lwd = 3, lty = 2, col = 1)
legend(x = 0.7, y = 0.3, legend = paste0("AUC = ", auc_rf))

```


## XGB

```{r}
# calculate ROC
perf_pred <- get_roc(fit_xgb_roc, df_test)
perf_roc_xgb <- perf_pred[[1]]
pred_roc_xgb <- perf_pred[[2]]


# take AUC 
auc_xgb <- round(unlist(slot(performance(pred_roc_xgb, measure = "auc"), "y.values")), 3)

# plot
plot(perf_roc_xgb, main = "XGB ROC curve", col = "steelblue", lwd = 3)
abline(a = 0, b = 1, lwd = 3, lty = 2, col = 1)
legend(x = 0.7, y = 0.3, legend = paste0("AUC = ", auc_xgb))

```


---
```{r}
save.image("workspace.RData")
```

