---
title: "Stroke prediction"
author: "AG"
date: "last update: `r format(Sys.Date(), format = '%d %B %Y')`"
output:
  html_document:
    toc: yes
    number_sections: true
    df_print: paged
    code_folding: hide
    fig_width: 10
    fig_height: 6
    theme: cerulean
    highlight: espresso
    toc_float: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F, cache = T)
library(tidyverse)
library(caret)
#library(GGally)
library(DMwR) # installed via archive: install.packages("/Path/to/DMwR_0.4.1.tar.gz", repos=NULL, type="source" )
library(ROCR)

```

# The data

In `smoking_status` Unknown should be changed to NA.

Also, it can be ordered: never < formerly < smokes

`ever_married` can be recoded as 0/1 in accordance with `heart_disease` and `hypertension`

Other predictors seem to be OK

```{r read}
df <- read_csv("data/healthcare-dataset-stroke-data.csv", col_types = "cfdfffffddcf", na = c("Unknown", "N/A"))

# if you set smoking_status to factor in col_types, na() won't work
df$smoking_status <- as_factor(df$smoking_status)
df$smoking_status <- fct_relevel(df$smoking_status, c("never smoked", "formerly smoked", "smokes"))

# married
df$ever_married <- factor(if_else(df$ever_married == "Yes", 1, 0))

# for models working properly
df$stroke <- factor(ifelse(df$stroke == 1, "yes", "no"), levels = c("no", "yes"))

df
```

## Descriptive statistics

Skip id column

```{r describe}
df$id <- NULL
skimr::skim(df)
```

Target 'stroke' is imbalanced!

'smoking_status' completeness rate 0.7

## How many smoking_status in each target class?

```{r}
df %>% group_by(stroke, smoking_status) %>% summarise(N=n())
```


BMI's complete rate 0.96

## How many skipped BMI in each target class?

```{r}
df %>% filter(is.na(bmi)) %>% group_by(stroke) %>% summarise(N=n())
```


One 'Other' gender to be removed

```{r filter}
df <- df %>% filter(gender != "Other")
```


## EDA

### Overview: a pairs plot

```{r pairs, fig.width=10, fig.height=10, warning=FALSE, message=FALSE, eval=FALSE}
GGally::ggpairs(df, aes(color = stroke, alpha = 0.2, dotsize = 0.02), 
        upper = list(continuous = GGally::wrap("cor", size = 2.5)),
        diag = list(continuous = "barDiag")) +
  scale_color_brewer(palette = "Set1", direction = -1) +
  scale_fill_brewer(palette = "Set1", direction = -1)
```

### In details

#### Stroke vs Age

```{r}
ggplot(df, aes(stroke, age)) +
  geom_boxplot(aes(fill = stroke), alpha = 0.5, varwidth = T, notch = T) +
  geom_violin(aes(fill = stroke), alpha = 0.5) +
  scale_fill_brewer(palette = "Set1", direction = -1) +
  xlab("")
```

**OBS!** There are observation with age much below 20 y.o., even close to 0!

These are very young kids or babies - should we even include them in the analysis?

If so, it will be prediction for adults. Also, stroke in kids probably has very different causes compared to stroke in adults/old folk.

#### Stroke vs Age + Gender

```{r box1}
ggplot(df, aes(stroke, age)) + 
  geom_violin(alpha=0.3) +
  geom_jitter(alpha=0.2, size=0.8, width = 0.15, height = 0.1, aes(color = gender)) + 
  geom_boxplot(alpha = 0.2) +
  scale_color_brewer(palette = "Set2", direction = -1)
```

#### Stroke vs Glucose

```{r}
ggplot(df, aes(stroke, avg_glucose_level)) +
  geom_boxplot(aes(fill = stroke), alpha = 0.5, varwidth = T, notch = T) +
  geom_violin(aes(fill = stroke), alpha = 0.5) +
  scale_fill_brewer(palette = "Set1", direction = -1) +
  xlab("")
```

This average glucose level is probably the results of fasting blood sugar test

If I correctly understand this [CDC information](https://www.cdc.gov/diabetes/basics/getting-tested.html#:~:text=Fasting%20Blood%20Sugar%20Test&text=A%20fasting%20blood%20sugar%20level,higher%20indicates%20you%20have%20diabetes.) on diabetes, values greater than 126 is evidence of diabetes.  But 250? Is it realistic?

#### Stroke vs BMI

```{r}
ggplot(df, aes(stroke, bmi)) +
  geom_boxplot(aes(fill = stroke), alpha = 0.5, varwidth = T, notch = T) +
  geom_violin(aes(fill = stroke), alpha = 0.5) +
  scale_fill_brewer(palette = "Set1", direction = -1) +
  xlab("")
```

BMI over 40 is the 3rd class of obesity - BMI over 75 should not exist at all.

Let's look at this weird points

#### Age vs BMI vs Glucose

```{r}
facet_names <- c("no" = "no stroke", "yes" = "stroke")

ggplot(df, aes(age, bmi)) +
  geom_point(color = "steelblue", alpha = 0.8, size = 0.5) +
  scale_fill_brewer(palette = "Set1", direction = -1) +
  facet_grid(rows = vars(stroke), labeller = as_labeller(facet_names)) +
  guides()
```

Patients with BMI over 75 are also very young. Suspicious.

#### Glucose vs Age + smoking

```{r}
ggplot(df, aes(age, avg_glucose_level)) +
  geom_point(aes(color = smoking_status), alpha = 0.6, size = 1) +
  scale_fill_brewer(palette = "Set1", direction = -1) +
  facet_grid(rows = vars(stroke), labeller = as_labeller(facet_names)) +
  guides()
```

**OBS!** Kids are mainly 'Unknown' smoking status; both target groups are divided into two clusters -- I'm curious why.

It is not gender, nor heart disease or any other factor we have in the data set.

#### Age vs Smoking

```{r}
ggplot(df, aes(smoking_status, age)) +
  geom_boxplot(aes(fill = stroke), alpha = 0.5, varwidth = T, notch = T) +
  scale_fill_brewer(palette = "Set1", direction = -1) +
  xlab("")
```

Kids are mainly non-smokers of course. Note the same two outliers of age below 20 in stroke-yes

#### Glucose vs BMI


```{r}
ggplot(df, aes(avg_glucose_level, bmi)) +
  geom_point(aes(color = age), alpha = 0.6, size = 1) +
  scale_fill_brewer(palette = "Set1", direction = -1) +
  facet_grid(rows = vars(stroke), labeller = as_labeller(facet_names)) +
  guides()
```

BMI outliers: super high BMI but super low glucose levels? How's that possible?

Can it be a bias introduced by testing protocol misuse? Like instead of measuring glucose *before* eating, in some samples it was done *after* eating?

#### Stroke vs Gender

```{r bar1}
gender <- df %>% group_by(stroke, gender) %>% summarize(N=n())

ggplot(gender, aes(stroke, N)) +
  geom_bar(aes(fill=gender), alpha = 0.8, stat = "identity", position = "fill") +
  scale_fill_brewer(palette = "Set2", direction = -1) +
  ylab("proportion")
```

Proportions in both stroke groups are roughly the same

#### Stroke vs Hypertension

```{r bar2}
hyptens <- df %>% group_by(stroke, hypertension) %>% summarize(N = n())

ggplot(hyptens, aes(stroke, N)) +
  geom_bar(aes(fill = hypertension), alpha = 0.8, stat = "identity", position = "fill") +
  scale_fill_brewer(palette = "Set2", direction = -1) +
  ylab("proportion")
```

Hypertension occurred more often in stroke-yes

#### Stroke vs Heart Disease

```{r}
heart <- df %>% group_by(stroke, heart_disease) %>% summarize(N=n())

ggplot(heart, aes(stroke, N)) +
  geom_bar(aes(fill = heart_disease), alpha = 0.8, stat = "identity", position = "fill") +
  scale_fill_brewer(palette = "Set2", direction = 1) +
  ylab("proportion")
```

#### Stroke vs Ever Married

```{r}
married <- df %>% group_by(stroke, ever_married) %>% summarize(N=n())

ggplot(married, aes(stroke, N)) +
  geom_bar(aes(fill = ever_married), alpha = 0.8, stat = "identity", position = "fill") +
  scale_fill_brewer(palette = "Set2", direction = -1) +
  ylab("proportion")
```

Marriage is bad :)

#### Stroke vs Work Type

```{r}
work <- df %>% group_by(stroke, work_type) %>% summarize(N=n())

ggplot(work, aes(stroke, N)) +
  geom_bar(aes(fill = work_type), alpha = 0.8, stat = "identity", position = "fill") +
  scale_fill_brewer(palette = "Set2", direction = 1) +
  ylab("proportion")
```

It's good to be a child

#### Stroke vs Residence Type

```{r}
residence <- df %>% group_by(stroke, Residence_type) %>% summarize(N=n())

ggplot(residence, aes(stroke, N)) +
  geom_bar(aes(fill = Residence_type), alpha = 0.8, stat = "identity", position = "fill") +
  scale_fill_brewer(palette = "Set2", direction = 1) +
  ylab("proportion")
```

#### Stroke vs Smoking

```{r}
smoking <- df %>% group_by(stroke, smoking_status) %>% summarize(N=n())

ggplot(smoking, aes(stroke, N)) +
  geom_bar(aes(fill = smoking_status), alpha = 0.8, stat = "identity", position = "fill") +
  scale_fill_brewer(palette = "Set2", direction = 1) +
  ylab("proportion")
```

#### Kids and Smoking

A lot of NAs in `smoking_status` comes from group 'Children' (see `work_type`). I can replace them with 'never smoked'.

```{r}
df %>% filter(work_type == "children") %>% 
  group_by(smoking_status) %>% 
  summarise(N = n(), 
            avg.age = mean(age), 
            max.age = max(age), 
            min.age = min(age))
```


### Conclusions

There are several suspicious outliers (like in BMI and glucose). I can safely remove BMI > 75, maybe even BMI > 60 (Remember that  BMI > 40 is the highest class of obesity).

What is less safe - it's removing non-adults (younger than 20 y.o.). On one hand they provide valid information (age is very important predictor), on the other hand their stroke cases are really sus and a lot of predictors do not have sense (or are obvious NAs) for non-adults (like smoking, marriage status, employment type, residence type etc.); model-based imputation of `smoking_status` in children doesn't have sense as well, I should rather replace with "never smoked".

Since, modelling using all predictors and observations has given me very moderate results (TPR = 1 comes with very high FPR and very low probability cutoff close to 0), I will try various trimming of the data.

## Trimming

Remove bmi > 60

```{r, cache=F}
df_trim <- df %>% filter(bmi <= 60 )

skimr::skim(df_trim)
```

BMI is complete, in total approx. 2000 observations are gone

## Imputation

Using package [mice](https://www.rdocumentation.org/packages/mice/versions/3.14.0/topics/mice)

It uses `polr` - proportional odds model - for `smoking_status` and `pmm` - predictive mean matching - for `bmi`

### Run imputation

```{r, eval = T, message=FALSE, cache=F}
library(mice)

imp_mice <- mice(df_trim)
df_imp <- complete(imp_mice)

```

Number of NAs in BMI: `r sum(is.na(df_imp$bmi))`

Number of NAs in Smoking: `r sum(is.na(df_imp$smoking_status))`

### Check distributions

#### BMI

```{r bmi.imp.plot}
bmi_imp_comp <- bind_rows(select(df_trim, bmi, stroke) %>% mutate(type = rep("original", nrow(df_trim))),
          select(df_imp, bmi, stroke) %>% mutate(type = rep("imputed", nrow(df_imp))))

ggplot(bmi_imp_comp, aes(bmi)) +
  geom_histogram(aes(fill = type), alpha = 0.8) +
  facet_grid(cols = vars(stroke))
```

Means have not changed, which is good, I suppose.

#### Smoking

```{r}
smoke_imp_comp <- bind_rows(select(df_trim, smoking_status, stroke) %>% mutate(type = rep("original", nrow(df_trim))),
          select(df_imp, smoking_status, stroke) %>% mutate(type = rep("imputed", nrow(df_imp))))

ggplot(smoke_imp_comp, aes(smoking_status)) +
  geom_bar(aes(fill=type), alpha=0.8, position="dodge") +
  facet_grid(cols = vars(stroke)) +
  xlab("")+
  theme(axis.text.x = element_text(angle=45, vjust = 0.5))
```

Counts increased proportionally in all *Smoking* groups


## Scaling & Normalization

Scale numeric features (including imputed BMI)

```{r scale, cache=F}
# use caret::preProcess()
# preProcValues <- preProcess(training, method = c("center", "scale"))

df_scaled <- df_imp %>% 
  select(avg_glucose_level, age, bmi) %>% 
  scale() %>% 
  data.frame()

```


## Make Dummies

I've decided to omit smoking_status completely - it won't be *dummified*

```{r dummy, cache=F}
# select vars
to_dum <- df_imp %>% select(gender, work_type, Residence_type, smoking_status)
# make an obj
dummies <- dummyVars(~ ., data = to_dum)
# apply it
df_dummy <- data.frame(predict(dummies, newdata = to_dum))

head(df_dummy)
```

## Join scaled and dummies and the rest

```{r join}
df_proc <- bind_cols(df_scaled, df_dummy, select(df_trim, hypertension, heart_disease, ever_married, stroke))
head(df_proc)
```

# Modelling

## Basic parameters

ROC-optimization is better when data is imbalanced

```{r tune1}
# for ROC
fit_ctrl_roc <- trainControl(## 5-fold CV
                           method = "repeatedcv",
                           number = 5,
                           repeats = 10, 
                           allowParallel = T,
                           classProbs = T,
                           summaryFunction = twoClassSummary)
```

## Split data

Imbalanced data - use SMOTE to create training data set, but not testing data set

```{r split}
set.seed(1234)
sample_set <- createDataPartition(y = df_proc$stroke, p = .75, list = FALSE)
df_train <- df_proc[sample_set,]
df_test <- df_proc[-sample_set,]

# DMwR::SMOTE for imbalanced data: over=225 and under=150 give me 1:1 ratio
df_train_smote <- SMOTE(stroke ~ ., data.frame(df_train), perc.over = 1725, perc.under = 106)

df_train_smote %>% group_by(stroke) %>% summarise(N=n())
```


# Random Forest

## Training and validation

```{r train.rf}
set.seed(121)

THREADS <- 6

library(doParallel)

cl <- makePSOCKcluster(THREADS)
registerDoParallel(cl)

fit_rf <- train(stroke ~ ., 
                 data = df_train_smote, 
                 metric = "ROC", 
                 method = "rf", 
                 trControl = fit_ctrl_roc,
                 tuneGrid = expand.grid(.mtry = c(7.5, 8.5, 9.0, 9.5)),
                 verbosity = 0,
                 ntree = 25,
                 nodesize = 1,
                 verbose = FALSE)

stopCluster(cl)

fit_rf

```


## Features importance

```{r featimp.rf}
imp_vars_rf <- varImp(fit_rf)

plot(imp_vars_rf, main = "Variable Importance with RF")

```


## Testing

### ROC & AUC

a function for roc-stuff

```{r roc.fun}
get_roc <- function(fit.obj, testing.df){
  pred_prob <- predict.train(fit.obj, newdata = testing.df, type = "prob")
  pred_roc <- prediction(predictions = pred_prob$yes, labels = testing.df$stroke)
  perf_roc <- performance(pred_roc, measure = "tpr", x.measure = "fpr")
  return(list(perf_roc, pred_roc))
}
```


### ROC-curve for ROC-optimized model

```{r roc.rf}
# calculate ROC
perf_pred <- get_roc(fit_rf, df_test)
perf_rf <- perf_pred[[1]]
pred_rf <- perf_pred[[2]]

# take AUC 
auc_rf <- round(unlist(slot(performance(pred_rf, measure = "auc"), "y.values")), 3)

# plot
plot(perf_rf, main = "Random Fores ROC curve", col = "steelblue", lwd = 3)
abline(a = 0, b = 1, lwd = 3, lty = 2, col = 1)
legend(x = 0.7, y = 0.3, legend = paste0("AUC = ", auc_rf))

```

So, we can adjust TPR/FPR cutoff to predict *all* stroke cases.

### TPR, FPR vs Probability cutoff

At which probability cut-off, you'll get TPR = 1.0?

```{r tpr.r}
# use pred_rf (pred_roc) object
plot(performance(pred_rf, measure = "tpr", x.measure = "cutoff"),
     col = "steelblue", 
     ylab = "Rate", 
     xlab = "Probability cutoff")

plot(performance(pred_rf, measure = "fpr", x.measure = "cutoff"), 
     add = T, col = "red")

legend(x = 0.65,y = 0.7, c("TPR (Recall)", "FPR (1-Spec)"), 
       lty = 1, col = c('steelblue', 'red'), bty = 'n', cex = 1, lwd = 2)

#abline(v = 0.02, lwd = 2, lty=6)

title("RF")
```


### Confusion matrix

Using desired cut-off: we want to maximize TPR (Sensitivity, Recall)!

According to the TPR/FPR plot (above) the optimal cutoff is.

```{r test.rf}
# predict probabilities
pred_prob_rf <- predict(fit_rf, newdata = df_test, type = "prob")

# choose your cut-off
cutoff = 0.01

# turn probabilities into classes
pred_class_rf <- ifelse(pred_prob_rf$yes > cutoff, "yes", "no")

pred_class_rf <- as.factor(pred_class_rf)

cm_rf <- confusionMatrix(data = pred_class_rf, 
                reference = df_test$stroke,
                mode = "everything",
                positive = "yes")

cm_rf
```


# Extreme Gradient Boosting: xgbTree

xgbTree has 7 parameters

## Training and validation

```{r train.xgb, eval=F}
set.seed(121)

# the best tune achieved by using tuneLength=4 is as the following:
# n_rounds = 50
# max_depth = 4
# eta = 0.3
# gamma = 0
# colsample_bytree = 0.6
# min_child_weight = 1
# subsample = 1

fit_xgb <- train(stroke ~ ., 
                 data = df_train_smote, 
                 method = "xgbTree",
                 metric = "ROC", 
                 trControl = fit_ctrl_roc,
                 tuneLength = 4,
                 nthreads = 16,
                 verbose = FALSE,
                 verbosity = 0)

fit_xgb$bestTune
```


## Features importance

```{r featimp.xgb, eval=F}
imp_vars_xgb <- varImp(fit_xgb)

plot(imp_vars_xgb, main = "Variable Importance with XGB")
```


## Testing

### ROC curve

```{r roc.xgb, eval=F}
# calculate ROC
perf_pred_xgb <- get_roc(fit_xgb, df_test)
perf_xgb <- perf_pred_xgb[[1]]
pred_xgb <- perf_pred_xgb[[2]]


# take AUC 
auc_xgb <- round(unlist(slot(performance(pred_xgb, measure = "auc"), "y.values")), 3)

# plot
plot(perf_xgb, main = "XGB ROC curve", col = "steelblue", lwd = 3)
abline(a = 0, b = 1, lwd = 3, lty = 2, col = 1)
legend(x = 0.7, y = 0.3, legend = paste0("AUC = ", auc_xgb))

```

### TPR v FPR

```{r tpr.xgb, eval=F}
# use pred_xgb object
plot(performance(pred_xgb, measure = "tpr", x.measure = "cutoff"),
     col = "steelblue", 
     ylab = "Rate", 
     xlab = "Probability cutoff")

plot(performance(pred_xgb, measure = "fpr", x.measure = "cutoff"), 
     add = T, col = "red")

legend(x = 0.6,y = 0.7, c("TPR (Recall)", "FPR (1-Spec)"), 
       lty = 1, col = c('steelblue', 'red'), bty = 'n', cex = 1, lwd = 2)

#abline(v = 0.1, lwd = 2, lty=6)

title("xgbTree")
```

### Confusion matrix

```{r cm.xgb, eval=F}
pred_prob_xgb <- predict(fit_xgb, newdata=df_test, type = "prob")

# choose your cut-off
cutoff = 0.12

# turn probabilities into classes
pred_class_xgb <- ifelse(pred_prob_xgb$yes > cutoff, "yes", "no")

pred_class_xgb <- as.factor(pred_class_xgb)

cm_xgb <- confusionMatrix(data = pred_class_xgb, 
                reference = df_test$stroke,
                mode = "everything",
                positive = "yes")

cm_xgb
```


# Save the workspace

---
```{r save, eval=T}
save.image("data/workspace.RData")
```

