---
title: "Matrix metalloproteinases (MMPs) English version"
author: "A.Guliaev"
date: "Feb 9, 2017"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
---

```{r setup, include=FALSE, echo=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
library(car)
library(ggplot2)
library(reshape)
library(knitr)
#library(gclus)
#library(RColorBrewer)
library(yarrr)
#library(beeswarm)
library(PMCMR)
library(psych)
library(stargazer, quietly = TRUE)

data.reduced <- read.csv("~/Dropbox/MCSC/pancreatology/data.reduced.EN.csv", na = "empty", stringsAsFactors = TRUE)
data.reduced.melt <- melt(data.reduced, id.vars = c("sex", "group", "cancer"))

```

## Descriptive statistics: data.reduced
```{r descript, echo=F, warning=F}
des <- describeBy(x = data.reduced.melt[,5], group = list(data.reduced.melt$group, data.reduced.melt$variable), mat=TRUE, digits = 2)
rownames(des) <- c()
kable(des[, c(-1, -4, -9, -10,-14, -15)])
```


## Scatter plot matrix: data.reduced
```{r matrices, echo=FALSE}
scatterplotMatrix(~MMP2+MMP9+TIMP2, data=data.reduced, main="", diagonal="qqplot")
```


## Scatter plot matrix: females
```{r matrices fem, echo=FALSE}
scatterplotMatrix(~MMP2+MMP9+TIMP2, data=data.reduced[data.reduced$sex=="fem",], main="", diagonal="qqplot")
```


## Scatter plot matrix: males
```{r matrices males, echo=FALSE}
scatterplotMatrix(~MMP2+MMP9+TIMP2, data=data.reduced[data.reduced$sex=="mal",], main="", diagonal="qqplot")
```


## Spearman correlation matrix: data.reduced
```{r corr all, echo=F, warning=F}
kable(cor(data.reduced[,4:6],method = "spearman"), digits=2)
kable(cor(data.reduced[data.reduced$sex=="fem",4:6],method = "spearman"), digits=2)
kable(cor(data.reduced[data.reduced$sex=="mal",4:6],method = "spearman"), digits=2)
```


### Spearman correlation: MMP2 ~ TIMP2
```{r corr mmp2~timp2, echo=F, warning=F}
cor.test(data.reduced$TIMP2, data.reduced$MMP2, method = "spearman")
# cor.test(data.reduced$TIMP2, data.reduced$MMP2, method = "pearson")
```


## Simple linear regression: data.reduced
```{r regr 1, echo=FALSE, warning=FALSE, message=FALSE}
fit <- lm(TIMP2~MMP2, data=data.reduced)
summary(fit)
#stargazer(fit, type="latex")
ggplot(data.reduced, aes(MMP2, TIMP2)) + geom_point(shape=1) + geom_smooth(method=lm)
```


## Kruskal-Wallis rank sum test: data.reduced
### MMP2~group
```{r kw test, echo=F, warning=FALSE}
kruskal.test(MMP2~group, data=data.reduced)
```

#### Boxplot

```{r mmp2 box, echo=F, warning=FALSE}
ggplot(data.reduced, aes(group, MMP2))+geom_boxplot()+geom_jitter()+xlab("")

```

#### Pirate plot

```{r, echo=F, warning=FALSE}
pirateplot(MMP2 ~ group, data.reduced, avg.line.fun = median,pal="grey", point.cex = .9, point.col = "black", bean.f.col="black", bean.f.o = .1, avg.line.col="black", bar.b.o = 0, bar.f.o = 0,inf.f.col = "grey", inf.f.o = 0.6,  inf.disp = "bean", xlab = "")
```


### MMP9~group -> p-value = 0.01201
```{r kw test 2, echo=F, warning=FALSE}
kruskal.test(MMP9~group, data=data.reduced)
```

#### Boxplot

```{r mmp9 box, echo=F, warning=F}
ggplot(data.reduced, aes(group, MMP9))+geom_boxplot()+geom_jitter()+xlab("")
```

#### Pirate plot

```{r mmp9 pirate, echo=F, warning=FALSE}
pirateplot(MMP9 ~ group, data.reduced, avg.line.fun = median,pal="grey", point.cex = .9, point.col = "black", bean.f.col="black", bean.f.o = .1, avg.line.col="black", bar.b.o = 0, bar.f.o = 0,inf.f.col = "grey", inf.f.o = 0.6,  inf.disp = "bean", xlab = "")
```


### MMP9~group Dunn's test
```{r mmp9 dunn, echo=F, warning=FALSE, message=FALSE}
kw.posthoc.mmp9<-posthoc.kruskal.dunn.test(data.reduced$MMP9, data.reduced$group)
summary(kw.posthoc.mmp9)
#kw.posthoc.mmp9$statistic
#kw.posthoc.mmp9$p.value
```


### TIMP2~group -> p-value = 0.03042
```{r kw test 3, echo=F, warning=FALSE, message=FALSE, results='asis'}
kruskal.test(TIMP2~group, data=data.reduced)
#stargazer(summary(w), type = "latex")
```

#### Boxplot

```{r timp2 box, echo=F, warning=F}
ggplot(data.reduced, aes(group, TIMP2))+geom_boxplot()+geom_jitter()+xlab("")
```

#### Pirate plot

```{r timp2 pirate, echo=F, warning=FALSE}
pirateplot(TIMP2 ~ group, data.reduced, avg.line.fun = median,pal="grey", point.cex = .9, point.col = "black", bean.f.col="black", bean.f.o = .1, avg.line.col="black", bar.b.o = 0, bar.f.o = 0,inf.f.col = "grey", inf.f.o = 0.6,  inf.disp = "bean", xlab = "")
```


### TIMP2~group Dunn's test
```{r timp2 dunn, echo=F}
kw.posthoc.timp2<-posthoc.kruskal.dunn.test(data.reduced$TIMP2, data.reduced$group)
summary(kw.posthoc.timp2)
```
