---
title: "Software Defects Detection"
author: "AG"
format: html
editor: source
---

## Data

```{r, message=FALSE}
library(tidyverse)
library(tidymodels)
library(ggpubr)
library(finetune)
library(bonsai)
library(embed)
library(themis)
library(baguette)

plot_validation_results <- function(dat, components) {
  # dat - data table
  # components - vector of components names to plot
  dat %>%
    select(-id) %>%
    select(all_of(c(components, "defects"))) %>% 
    # Create the scatterplot matrix
    ggplot(aes(x = .panel_x, y = .panel_y, color = defects, fill = defects)) +
    geom_point(alpha = 0.2, size = 1) +
    ggforce::geom_autodensity(alpha = .3) +
    ggforce::facet_matrix(vars(-defects), layer.diag = 2) + 
    scale_color_brewer(palette = "Dark2") + 
    scale_fill_brewer(palette = "Dark2")
}
```

Defects is boolean originally, I'll make it factor (1, 0)

```{r}
data_synth <- read_csv("data/train.csv.gz", show_col_types = F) %>% 
  mutate(defects = factor(if_else(defects, 1, 0), levels=c("1", "0")))

data_real <- read_csv("data/jm1.csv", show_col_types = F) %>% 
  mutate(defects = factor(if_else(defects, 1, 0), levels=c("1", "0")),
         id = c(nrow(data_synth): nrow(data_synth) + nrow(data_real))) %>% 
  relocate(id, .before = loc)


data_synth %>% head()
```

```{r}
data_real %>% head()
```


```{r}
setdiff(names(data_synth), names(data_real))
```


```{r}
data_synth %>% skimr::skim()
```

```{r}
data_real %>% skimr::skim()
```

### Join the data

```{r}
data_all <- bind_rows(data_synth, data_real) %>% 
  drop_na(names(data_real)[18:22])

data_all %>% skimr::skim()
```


## EDA

### Histograms as is

```{r, fig.height=10}
plot_hist <- function(varname, df){
  df %>% 
    ggplot(aes(.data[[varname]])) +
    geom_histogram(aes(fill=defects), bins=200) +
    geom_rug()
}

hist_list <- map(names(data_all)[2:22], ~ plot_hist(., data_all))

ggarrange(plotlist = hist_list, nrow = 7, ncol = 3, common.legend = T)
```

Let's apply some transformations

### Histograms after transform

```{r, fig.height=10}
rec_trans <- recipe(defects ~ ., data = data_all) %>% 
  update_role(id, new_role = "ID") %>% 
  step_zv(all_predictors()) %>% 
  step_log(all_numeric_predictors(), offset = 1) %>% 
  step_normalize(all_numeric_predictors())

data_log <- rec_trans %>% prep() %>% juice()

hist_list <- map(names(data_log)[2:22], ~ plot_hist(., data_log))

ggarrange(plotlist = hist_list, nrow = 7, ncol = 3, common.legend = T)
```

### Boxplots

```{r, fig.height=12}
plot_box <- function(varname, df){
  df %>% 
    ggplot(aes(defects, .data[[varname]])) +
    geom_boxplot(aes(fill=defects), notch = T) +
    xlab("")
}

box_list <- map(names(data_log)[2:22], ~ plot_box(., data_log))

ggarrange(plotlist = box_list, nrow = 7, ncol = 3, common.legend = T)

```

### Corrplot

```{r, fig.width=8}
library(ggcorrplot)

corr_matrix <- round(cor(data_all[,-c(1, 23)]), 1) 

ggcorrplot(corr_matrix,
           hc.order = TRUE,
           type = "upper",
           lab = TRUE)
```


### UMAP

In the recipes below I do down-sampling

#### Non-supervised

```{r, fig.height=8, fig.width=8}
umap_yj_recipe <- 
  recipe(defects ~ ., data = data_synth) %>%
  update_role(id, new_role = "ID") %>%
  step_nzv(all_predictors()) %>% 
  step_log(all_numeric_predictors(), offset = 1) %>% 
  step_normalize(all_numeric_predictors()) %>%
  step_downsample(defects, under_ratio = 1, seed = 99) %>% 
  step_umap(
    all_numeric_predictors(),
    num_comp = 10,
    min_dist = 0.8,
    neighbors = 25
  ) %>% 
  step_normalize(all_numeric_predictors())

umap_results <- umap_yj_recipe %>%
  prep() %>% 
  juice() 


umap_results %>% 
  slice_sample(n = 1000, by = defects) %>% 
  plot_validation_results(c("UMAP01", "UMAP02", "UMAP03", "UMAP04")) +
  ggtitle("UMAP unsupervised")
```

#### Supervised

```{r, fig.height=8, fig.width=8}
umap_sup_recipe <-
  recipe(defects ~ ., data = data_synth) %>%
  update_role(id, new_role = "ID") %>%
  step_nzv(all_predictors()) %>%
  step_log(all_numeric_predictors(), offset = 1) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_downsample(defects, under_ratio = 1, seed = 99) %>% 
  step_umap(
    all_numeric_predictors(),
    outcome = "defects",
    num_comp = 10,
    min_dist = 0.8,
    neighbors = 25
  )

umap_sup_results <- umap_sup_recipe %>%
  prep() %>% 
  juice() 


umap_sup_results %>% 
  slice_sample(n = 1000, by = defects) %>% 
  plot_validation_results(c("UMAP01", "UMAP02", "UMAP03", "UMAP04")) +
  ggtitle("UMAP supervised")
```

### PCA

```{r, fig.height=8, fig.width=8}
rec_pca <- 
  recipe(defects ~ ., data = data_synth) %>%
  update_role(id, new_role = "ID") %>%
  step_nzv(all_predictors()) %>%
  step_log(all_numeric_predictors(), offset = 1) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_downsample(defects, under_ratio = 1, seed = 99) %>% 
  step_pca(all_numeric_predictors(), num_comp = 8) %>% 
  step_normalize(all_numeric_predictors())

rec_pca %>% 
  prep() %>% 
  juice() %>% 
  slice_sample(n = 1000, by = defects) %>% 
  plot_validation_results(c("PC1", "PC2", "PC3", "PC4", "PC5")) +
  ggtitle("Principal Component Analysis")
```

```{r, fig.height=8, fig.width=8}
rec_kpca <- 
  recipe(defects ~ ., data = data_synth %>% slice_sample(n = 1000, by = defects)) %>%
  update_role(id, new_role = "ID") %>%
  step_nzv(all_predictors()) %>%
  step_log(all_numeric_predictors(), offset = 1) %>%
  step_downsample(defects, under_ratio = 1, seed = 99) %>% 
  step_normalize(all_numeric_predictors()) %>%
  step_kpca_poly(all_numeric_predictors(), num_comp = 6) %>% 
  step_normalize(all_numeric_predictors())

kpca_results <- rec_kpca %>% 
  prep() %>% 
  juice()

kpca_results %>% 
  plot_validation_results(c("kPC1", "kPC2", "kPC3", "kPC4")) +
  ggtitle("Kernel PCA")

```

## Data split, CV folds, metric, basic recipe

```{r}
library(themis)

# split
set.seed(124)

data_split <- initial_split(data_all, prop = 0.99, strata = defects)

df_train <- training(data_split)
df_test <- testing(data_split)

# folds
cv_folds <- vfold_cv(df_train, strata = "defects", v = 10, repeats = 5) 
# metrics for imbalanced classes
metrics <- metric_set(roc_auc)

# recipe
rec_base <-
  recipe(defects ~ ., data = df_train) %>%
  update_role(id, new_role = "ID") %>%
  step_nzv(all_predictors()) %>%
  step_log(all_numeric_predictors(), offset = 1) %>%
  step_normalize(all_numeric_predictors()) %>% 
  step_downsample(defects, under_ratio = 1, seed = 100)

rec_base %>% prep() %>% juice() %>% group_by(defects) %>% count()
```

## LR basic

```{r}
lr_spec <- logistic_reg(penalty = tune(),
                        mixture = 1) %>%
  set_engine("glmnet")

lr_wf <- workflow(preprocessor = rec_base,
                  spec = lr_spec)

lr_res <- lr_wf %>% 
  tune_grid(
    grid = 20,
    resamples = cv_folds,
    control = control_grid(save_pred = TRUE,
                           save_workflow = FALSE),
    metrics = metric_set(roc_auc)
  )

lr_res %>% show_best()

lr_best_param <- select_best(lr_res)
lr_model_final <- finalize_model(lr_spec, lr_best_param)
# should data set be processed here?
# proc and not-proc fit objects are not identical
lr_fit_proc <- fit(lr_model_final, defects ~ ., rec_base %>% prep() %>% juice())

saveRDS(lr_fit_proc, "models/lr_base_fit_proc.rds")

lr_res %>% show_best()
```

### Prediction

Empirically demonstrated: processed fit and processed test set give the best public test score

```{r}
data_test <- read_csv("data/test.csv.gz", show_col_types = F)

data_test_proc <- rec_base %>% prep() %>% bake(new_data = data_test)

test_prediction <- predict(lr_fit_proc, data_test_proc) %>% 
  mutate(id = data_test$id, .before = .pred_class) %>% 
  rename("defects" = .pred_class)

write_csv(x =  test_prediction, file = "submissions/test_prediction_lr_base.csv")
```


## GBT Race

```{r}
gbt_spec <- boost_tree(
  mtry = tune(),
  trees = tune(),
  tree_depth = tune(),
  learn_rate = tune(),
  min_n = tune(),
  loss_reduction = tune(), 
  sample_size = tune(), 
  stop_iter = tune()
) %>%
  set_engine("lightgbm", num.threads=8) %>%
  set_mode("classification") %>%
  translate()

set.seed(124)

gbt_wf_base <-
  workflow(preprocessor = rec_base, 
           spec = gbt_spec) 

param_set <- extract_parameter_set_dials(gbt_wf_base) %>%
    finalize(x = df_train %>% select(-defects))

gbt_base_wl_res <- gbt_wf_base %>%
  tune_race_win_loss(
    param_info = param_set,
    resamples = cv_folds,
    grid = 30,
    metrics = metric_set(roc_auc),
    control = control_race(
      verbose_elim = TRUE,
      save_pred = TRUE,
      save_workflow = FALSE,
      burn_in = 10
    )
  )

show_best(gbt_base_wl_res)
```

### Prediction

```{r}
gbt_best_param <- select_best(lgb_base_wl_res)
gbt_model_final <- finalize_model(gbt_spec, gbt_best_param)
# should data set be processed here?
# proc and not-proc fit objects are not identical
gbt_fit_proc <- fit(gbt_model_final, defects ~ ., rec_base %>% prep() %>% juice())

saveRDS(gbt_fit_proc, "models/gbt_base_fit_proc.rds")

test_prediction <- predict(gbt_fit_proc, data_test_proc) %>% 
  mutate(id = data_test$id, .before = .pred_class) %>% 
  rename("defects" = .pred_class)

write_csv(x =  test_prediction, file = "submissions/test_prediction_gbt_base.csv")
```

## GBT w PCA

```{r}
rec_pca <-
  recipe(defects ~ ., data = df_train) %>%
  update_role(id, new_role = "ID") %>%
  step_nzv(all_predictors()) %>%
  step_log(all_numeric_predictors(), offset = 1) %>%
  step_normalize(all_numeric_predictors()) %>% 
  step_downsample(defects, under_ratio = 1, seed = 100) %>% 
  step_pca(all_numeric_predictors(), num_comp = tune()) %>% 
  step_normalize(all_numeric_predictors())

set.seed(124)

gbt_wf_pca <-
  workflow(preprocessor = rec_pca, 
           spec = gbt_spec) 

param_set <- extract_parameter_set_dials(gbt_wf_pca) %>%
    finalize(x = df_train %>% select(-defects))

gbt_pca_wl_res <- gbt_wf_pca %>%
  tune_race_win_loss(
    param_info = param_set,
    resamples = cv_folds,
    grid = 30,
    metrics = metric_set(roc_auc),
    control = control_race(
      verbose_elim = TRUE,
      save_pred = TRUE,
      save_workflow = FALSE,
      burn_in = 10
    )
  )

show_best(gbt_pca_wl_res)
```

### Prediction

```{r}
data_test <- read_csv("data/test.csv.gz", show_col_types = F)

rec_pca$steps[[5]] <- update(rec_pca$steps[[5]], num_comp = 1)

data_test_proc <- rec_pca %>% prep() %>% bake(new_data = data_test)

gbt_best_param <- select_best(gbt_pca_wl_res)
gbt_model_final <- finalize_model(gbt_spec, gbt_best_param)
# should data set be processed here?
# proc and not-proc fit objects are not identical
gbt_fit_proc <- fit(gbt_model_final, defects ~ ., rec_pca %>% prep() %>% juice())

saveRDS(gbt_fit_proc, "models/gbt_pca_fit_proc.rds")

test_prediction <- predict(gbt_fit_proc, data_test_proc) %>% 
  mutate(id = data_test$id, .before = .pred_class) %>% 
  rename("defects" = .pred_class)

write_csv(x =  test_prediction, file = "submissions/test_prediction_gbt_pca.csv")
```


## GBT with grid search and Bayes opt

```{r}
gbt_spec <- boost_tree(
  trees = tune(),
  tree_depth = tune(),
  learn_rate = tune(),
  min_n = tune(),
  loss_reduction = tune(), 
  sample_size = tune(), 
  stop_iter = tune()
) %>%
  set_engine("lightgbm", num.threads=8) %>%
  set_mode("classification") %>%
  translate()

set.seed(124)

gbt_wf_base <-
  workflow(preprocessor = rec_base, 
           spec = gbt_spec) 

param_set <- extract_parameter_set_dials(gbt_wf_base) %>%
    finalize(x = df_train %>% select(-defects))

gbt_base_res <- gbt_wf_base %>%
  tune_grid(
    grid = 30,
    param_info = param_set,
    resamples = cv_folds,
    control = control_grid(save_pred = TRUE,
                           save_workflow = FALSE),
    metrics = metric_set(roc_auc)
  )



# Bayes
gbt_base_bres <-
      tune_bayes(
        gbt_wf_base,
        resamples = cv_folds,
        initial = gbt_base_res,
        iter = 30,
        metrics = metric_set(roc_auc),
        param_info = param_set,
        control = control_bayes(
          no_improve = 20,
          save_pred = TRUE,
          verbose = FALSE,
          save_workflow = FALSE
        )
      )


show_best(gbt_base_res)
show_best(gbt_base_bres)
```


## MLP BAG

```{r}
rec_nocorr <- 
  recipe(defects ~ ., data = df_train) %>%
  update_role(id, new_role = "ID") %>%
  step_nzv(all_predictors()) %>%
  step_log(all_numeric_predictors(), offset = 1) %>%
  step_normalize(all_numeric_predictors()) %>% 
  step_downsample(defects, under_ratio = 1, seed = 100) %>% 
  step_corr(all_predictors(), threshold = tune("corr_tune"))

mlp_spec <- bag_mlp(
      hidden_units = tune(),
      penalty = tune(),
      epochs = tune()) %>%
      set_engine("nnet", num.threads = 8) %>%
      set_mode("classification") %>%
      translate()

mlp_wf <-
  workflow(preprocessor = rec_nocorr, 
           spec = mlp_spec)

mlp_res <- mlp_wf %>% 
  tune_grid(
    grid = 30,
    resamples = cv_folds,
    control = control_grid(save_pred = TRUE,
                           save_workflow = FALSE),
    metrics = metric_set(roc_auc)
  )

mlp_res %>% show_best()
```

### Prediction

```{r}
data_test <- read_csv("data/test.csv.gz", show_col_types = F)

mlp_best_param <- select_best(mlp_res)
mlp_model_final <- finalize_model(mlp_spec, mlp_best_param)

rec_nocorr$steps[[5]] <- update(rec_nocorr$steps[[5]], threshold = mlp_best_param$corr_tune)

data_test_proc <- rec_nocorr %>% prep() %>% bake(new_data = data_test)

mlp_fit_proc <- fit(mlp_model_final, defects ~ ., rec_nocorr %>% prep() %>% juice())

saveRDS(mlp_fit_proc, "models/mlp_fit_proc.rds")

test_prediction <- predict(mlp_fit_proc, data_test_proc) %>% 
  mutate(id = data_test$id, .before = .pred_class) %>% 
  rename("defects" = .pred_class)

write_csv(x =  test_prediction, file = "submissions/test_prediction_mlp_bag.csv")
```

## RF basic

```{r}
rf_spec <-
  rand_forest(mtry = tune(), min_n = tune()) %>%
  set_engine('ranger', num.threads=8) %>%
  set_mode('classification')

rf_wf <- workflow(preprocessor = rec_base, spec = rf_spec)

rf_res <- rf_wf %>% 
  tune_grid(
    grid = 30,
    resamples = cv_folds,
    control = control_grid(save_pred = TRUE,
                           save_workflow = FALSE),
    metrics = metric_set(roc_auc)
  )

rf_res %>% show_best()
```

### Prediction

```{r}
data_test <- read_csv("data/test.csv.gz", show_col_types = F)

data_test_proc <- rec_base %>% prep() %>% bake(new_data = data_test)

rf_best_param <- select_best(rf_res)
rf_model_final <- finalize_model(rf_spec, rf_best_param)
# should data set be processed here?
# proc and not-proc fit objects are not identical
rf_fit_proc <- fit(rf_model_final, defects ~ ., rec_base %>% prep() %>% juice())

saveRDS(rf_fit_proc, "models/rf_base_fit_proc.rds")
```

## Stacks

### Specs

```{r}
library(stacks)

ctrl_grid <- control_stack_grid()
ctrl_res <- control_stack_resamples()

# model definitions
boost_tree_lgboost_spec <-
  boost_tree(
    tree_depth = tune(),
    trees = tune(),
    learn_rate = tune(),
    min_n = tune(),
    loss_reduction = tune(),
    sample_size = tune(),
    stop_iter = tune()
  ) %>%
  set_engine('lightgbm', num.threads=8) %>%
  set_mode('classification') %>%
  translate()

logistic_reg_glmnet_spec <-
  logistic_reg(penalty = tune(), mixture = tune()) %>%
  set_engine('glmnet', num.threads=8)

mlp_nnet_spec <-
  mlp(hidden_units = tune(), penalty = tune(), epochs = tune()) %>%
  set_engine('nnet', num.threads=8) %>%
  set_mode('classification')

# nearest_neighbor_kknn_spec <-
#   nearest_neighbor(
#     neighbors = tune(),
#     weight_func = tune(),
#     dist_power = tune()
#   ) %>%
#   set_engine('kknn', num.threads=8) %>%
#   set_mode('classification')
# 
# svm_poly_kernlab_spec <-
#   svm_poly(
#     cost = tune(),
#     degree = tune(),
#     scale_factor = tune(),
#     margin = tune()
#   ) %>%
#   set_engine('kernlab', num.threads=8) %>%
#   set_mode('classification')
```


### Recipes

```{r, fig.height=6}
# recipes
rec_base <-
  recipe(defects ~ ., data = df_train) %>%
  update_role(id, new_role = "ID") %>%
  step_nzv(all_predictors()) %>%
  step_log(all_numeric_predictors(), offset = 1) %>%
  step_normalize(all_numeric_predictors()) %>% 
  step_downsample(defects, under_ratio = 1, seed = 100)

rec_corr_tune <- 
  recipe(defects ~ ., data = df_train) %>%
  update_role(id, new_role = "ID") %>%
  step_nzv(all_predictors()) %>%
  step_log(all_numeric_predictors(), offset = 1) %>%
  step_normalize(all_numeric_predictors()) %>% 
  step_downsample(defects, under_ratio = 1, seed = 100) %>% 
  step_corr(all_predictors(), threshold = tune("corr_tune"))

rec_corr <-
  recipe(defects ~ ., data = df_train) %>%
  update_role(id, new_role = "ID") %>%
  step_nzv(all_predictors()) %>%
  step_log(all_numeric_predictors(), offset = 1) %>%
  step_normalize(all_numeric_predictors()) %>% 
  step_downsample(defects, under_ratio = 1, seed = 100) %>% 
  step_corr(all_predictors(), threshold = 0.9)

df_train_corr <- rec_corr %>% prep() %>% juice() 

corr_matrix <- round(cor(df_train_corr[,-c(1, ncol(df_train_corr))]), 1) 

# corr threshold 0.8 leaves only 6 predictors!
ggcorrplot(corr_matrix,
           hc.order = TRUE,
           type = "upper",
           lab = TRUE)
```

### Workflows

```{r}
lr_wf <- workflow(rec_corr, logistic_reg_glmnet_spec)

mlp_wf <- workflow(rec_corr, mlp_nnet_spec)

gbt_wf <- workflow(rec_corr, boost_tree_lgboost_spec)

```


### Tuning & validation

Via repeated CV, grid of size 20

```{r}

# param_set <- extract_parameter_set_dials(gbt_wf) %>%
#     finalize(x = df_train %>% select(-defects))

gbt_res <- gbt_wf %>%
  tune_grid(
    grid = 20,
    #param_info = param_set,
    resamples = cv_folds,
    control = ctrl_grid,
    metrics = metric_set(roc_auc)
  )

lr_res <- lr_wf %>% 
  tune_grid(
    grid = 20,
    resamples = cv_folds,
    control = ctrl_grid,
    metrics = metric_set(roc_auc)
  )

mlp_res <- mlp_wf %>%
  tune_grid(
    grid = 20,
    resamples = cv_folds,
    control = ctrl_grid,
    metrics = metric_set(roc_auc)
  )


saveRDS(lr_res, "models/lr_res_stacks.rds")
saveRDS(mlp_res, "models/mlp_res_stacks.rds")
saveRDS(gbt_res, "models/gbt_res_stacks.rds")

#lr_res <- readRDS("models/lr_res_stacks.rds")
#mlp_res <- readRDS("models/mlp_res_stacks.rds")
# gbt_res <- readRDS("models/gbt_res_stacks.rds")
```

```{r}
gbt_res %>% show_best()
```

```{r}
mlp_res %>% show_best()
```


### Stacking

```{r}
model_stacks <- stacks() %>% 
  add_candidates(lr_res) %>% 
  add_candidates(mlp_res) %>% 
  add_candidates(gbt_res)

as_tibble(model_stacks)
```

```{r}
model_stack_blend <- model_stacks %>% 
  blend_predictions(penalty = 10^(-6:-1),
                    mixture = 1)

autoplot(model_stack_blend)
```

```{r}
autoplot(model_stack_blend, type = "members")
```

```{r}
autoplot(model_stack_blend, type="weights")
```

### Look at params

```{r}
model_stack_fit <- model_stack_blend %>% fit_members()

collect_parameters(model_stack_fit, "gbt_res")
```

### Prediction

```{r}
data_test <- read_csv("data/test.csv.gz", show_col_types = F)

data_test_proc <- rec_corr %>% prep() %>% bake(new_data = data_test)

# Don't process the test data here
test_prediction <- predict(model_stack_fit, data_test) %>% 
  mutate(id = data_test$id, .before = .pred_class) %>% 
  rename("defects" = .pred_class)

write_csv(x =  test_prediction, file = "submissions/test_prediction_stacks3_corr.csv")
```


on my test

```{r}
my_test_pred <-
  df_test %>%
  bind_cols(predict(model_stack_fit, ., type = "prob"))

yardstick::roc_auc(
  my_test_pred,
  truth = defects,
  contains(".pred_1")
  )
```

Then on the public test data, this model stack gives ROC AUC = 0.66

Why?

## GBT with big grid

```{r}
gbt_spec <- boost_tree(
  trees = tune(),
  tree_depth = tune(),
  learn_rate = tune(),
  min_n = tune(),
  loss_reduction = tune(),
  sample_size = tune(),
  stop_iter = tune()
) %>%
  set_engine("lightgbm", num.threads = 8) %>%
  set_mode("classification") %>%
  translate()

set.seed(124)

gbt_wf_base <-
  workflow(preprocessor = rec_corr,
           spec = gbt_spec)

param_set <- extract_parameter_set_dials(gbt_wf_base) %>%
  finalize(x = df_train %>% select(-defects))

# Bayes
gbt_base_bres <-
  tune_bayes(
    gbt_wf_base,
    resamples = cv_folds,
    initial = 60,
    iter = 30,
    metrics = metric_set(roc_auc),
    param_info = param_set,
    control = control_bayes(
      no_improve = 30,
      save_pred = TRUE,
      verbose = FALSE,
      save_workflow = FALSE
    )
  )

show_best(gbt_base_bres)
```

```{r}
data_test <- read_csv("data/test.csv.gz", show_col_types = F)

gbt_best_param <- select_best(gbt_base_bres)
gbt_model_final <- finalize_model(gbt_spec, gbt_best_param)

data_test_proc <- rec_corr %>% prep() %>% bake(new_data = data_test)

gbt_fit_proc <- fit(gbt_model_final, defects ~ ., rec_corr %>% prep() %>% juice())

saveRDS(gbt_fit_proc, "models/gbt_B2_fit_proc.rds")

test_prediction <- predict(gbt_fit_proc, data_test_proc) %>% 
  mutate(id = data_test$id, .before = .pred_class) %>% 
  rename("defects" = .pred_class)

write_csv(x =  test_prediction, file = "submissions/test_prediction_gbt_B2.csv")
```


