---
title: "Health Outcome of Horses"
author: "AG"
format: html
editor: source
---

# Libs

```{r, message=FALSE}
library(tidyverse)
library(tidymodels)
library(embed)
library(skimr)
library(themis)
```


# Data

```{r}
data_synth <- read_csv("data/train.csv", show_col_types = F) %>% 
  mutate(across(is_character, ~ as_factor(.x)),
         hospital_number = as_factor(hospital_number))

data_synth %>% 
  skim()
```

```{r}
data_orig <- read_csv("data/horse_original.csv", show_col_types = F) %>% 
  mutate(across(is_character, ~ as_factor(.x)),
         hospital_number = as_factor(hospital_number))

data_orig %>% 
  skim()
```

# EDA

## Boxplots


```{r, fig.width=10, fig.height=10}
make_box <- function(df, column_name) {
  ggplot(df, aes(outcome, .data[[column_name]])) +
    geom_boxplot(varwidth = T) +
    geom_violin(alpha = 0.1) +
    xlab("")
}

box_list <- map(colnames(select(data_synth, where(is.numeric)))[2:11],
                ~ make_box(data_synth, .x))

ggpubr::ggarrange(plotlist =  box_list, ncol = 3, nrow = 4)
```

### Closer look at the lesions

```{r}
data_synth %>% 
  ggplot(aes(lesion_1)) +
  geom_histogram(aes(fill=outcome)) +
  facet_grid(rows = vars(outcome))
```

```{r}
data_synth %>% 
  ggplot(aes(lesion_2)) +
  geom_histogram(aes(fill=outcome), bins=60) +
  facet_grid(rows = vars(outcome))
```

```{r}
data_synth %>% 
  ggplot(aes(lesion_3)) +
  geom_histogram(aes(fill=outcome), bins=60) +
  facet_grid(rows = vars(outcome))
```

```{r}
data_synth %>% pull(lesion_1) %>% as_factor() %>% summary
```

```{r}
data_synth %>% pull(lesion_2) %>% as_factor() %>% summary
```

```{r}
data_synth %>% pull(lesion_3) %>% as_factor() %>% summary
```

```{r}
data_synth %>% group_by(outcome) %>% select(contains("lesion_"))
```

## Proportions bar plots

Excluding hospital number

```{r, fig.width=10, fig.height=10}
make_prop <- function(df, column_name, pos = "fill") {
  ggplot(df, aes(.data[[column_name]])) +
    geom_bar(aes(fill = outcome), position = pos) +
    ylab("") + theme(axis.text.x = element_text(
      angle = 45,
      vjust = 0.5,
      hjust = 0.5,
      size = 8
    )) +
    scale_fill_brewer(palette="Dark2")
}


prop_list <-
  map(colnames(select(data_synth, where(is.factor)))[c(1, 2, 4:17)],
      ~ make_prop(data_synth, .x))

ggpubr::ggarrange(plotlist =  prop_list, ncol = 4, nrow = 4, common.legend = T)
```

Hospital number

```{r, fig.width=14, fig.height=6}
make_prop(data_synth, "hospital_number")
```

### Frequencies of categorical variables

Excluding hospital number

```{r, fig.width=10, fig.height=10}
prop_list <-
  map(colnames(select(data_synth, where(is.factor)))[c(1, 2, 4:17)],
      ~ make_prop(data_synth, .x, pos = "stack"))

ggpubr::ggarrange(plotlist =  prop_list, ncol = 4, nrow = 4, common.legend = T)
```


## UMAP

```{r, fig.width=6, fig.height=6}
library(ggforce)

plot_dimred_results <- function(dat, color_by = outcome) {
  dat %>%
    select(-id) %>% 
    select(contains(c("01", "02", "03", "04")), {{color_by}}) %>% 
    # Create the scatterplot matrix
    ggplot(aes(x = .panel_x, y = .panel_y, color = {{color_by}})) +
    geom_point(alpha = 0.4, size = 1) +
    geom_autodensity(alpha = .3) +
    facet_matrix(vars(-{{color_by}}), layer.diag = 2) +
    scale_color_brewer(palette="Dark2")
}

umap_recipe <- recipe(outcome ~ ., data = data_synth) %>%
  update_role(id, new_role = "ID") %>% 
  step_lencode_glm(hospital_number, outcome = "outcome") %>% 
  step_other(all_nominal()) %>% 
  step_nzv(all_predictors()) %>%
  step_select(-c(hospital_number, contains("lesion_"))) %>% 
  step_dummy(all_nominal_predictors(), one_hot = T) %>% 
  step_normalize(all_numeric_predictors()) %>%
  step_YeoJohnson(all_numeric_predictors()) %>% 
  step_normalize(all_numeric_predictors()) %>% 
  step_umap(
    all_numeric_predictors(),
    num_comp = 10,
    min_dist = 0.01,
    neighbors = 25
  )

umap_results <- prep(umap_recipe, retain = TRUE)

umap_results$template %>% 
  plot_dimred_results() +
  ggtitle("UMAP")
```

```{r, fig.width=6, fig.height=6}
umap_sup_recipe <- recipe(outcome ~ ., data = data_synth) %>%
  update_role(id, new_role = "ID") %>% 
  step_lencode_glm(hospital_number, outcome = "outcome") %>% 
  step_other(all_nominal()) %>% 
  step_nzv(all_predictors()) %>%
  step_select(-c(hospital_number, contains("lesion_"))) %>% 
  step_dummy(all_nominal_predictors(), one_hot = T) %>% 
  step_normalize(all_numeric_predictors()) %>%
  step_YeoJohnson(all_numeric_predictors()) %>% 
  step_normalize(all_numeric_predictors()) %>% 
  step_umap(
    all_numeric_predictors(),
    num_comp = 10,
    min_dist = 0.01,
    neighbors = 25, 
    outcome = "outcome"
  )

umap_sup_results <- prep(umap_sup_recipe, retain = TRUE)

umap_sup_results$template %>% 
  plot_dimred_results() +
  ggtitle("UMAP supervised")
```

## PCA

```{r}
pca_recipe <- recipe(outcome ~ ., data = data_synth) %>%
  update_role(id, new_role = "ID") %>%
  update_role(outcome, new_role = "outcome") %>% 
  step_dummy(all_nominal_predictors(), one_hot = T) %>% 
  step_zv(all_predictors()) %>% 
  step_nzv(all_predictors()) %>% 
  step_normalize(all_numeric_predictors()) %>%
  step_YeoJohnson(all_numeric_predictors()) %>% 
  step_normalize(all_numeric_predictors()) %>%
  step_pca(all_predictors(), threshold = .8) %>% 
  step_normalize(all_numeric_predictors())

pca_results <- prep(pca_recipe, retain = TRUE)

pca_results$template %>% dim()
```

```{r, fig.width=8, fig.height=8}
pca_results$template %>% 
  plot_dimred_results() +
  ggtitle("PCA")
```

# Modelling on synth data

## Split (non-stratified)

```{r}
set.seed(124)

data_split <- initial_split(data_synth, prop = 0.99)

df_train <- training(data_split)
df_test <- testing(data_split)

cv_folds <- vfold_cv(df_train, v = 10, repeats = 5) 

f1_micro <- metric_tweak("f1_micro", f_meas, estimator = "micro")
```

## Recipes

```{r}
# my basic recipe with YJ transform
recipe_base <- recipe(outcome ~ ., data = df_train) %>%
  update_role(id, new_role = "ID") %>%
  step_nzv(all_nominal_predictors()) %>% 
  step_dummy(all_nominal_predictors(), one_hot = T) %>%
  step_nzv(all_predictors()) %>%
  step_YeoJohnson(all_numeric_predictors()) %>%
  step_normalize(all_numeric_predictors())

# basic recipe with no lesions and no hospital number
recipe_base_no_lesions_hosp_number <- recipe(outcome ~ ., data = df_train) %>%
  update_role(id, new_role = "ID") %>%
  step_select(-c(hospital_number, contains("lesion_"))) %>% 
  step_nzv(all_nominal_predictors()) %>% 
  step_dummy(all_nominal_predictors(), one_hot = T) %>%
  step_nzv(all_predictors()) %>%
  step_YeoJohnson(all_numeric_predictors()) %>%
  step_normalize(all_numeric_predictors())

# base recipe with SMOTE
recipe_base_smote <- recipe_base %>% 
  step_smote(outcome, over_ratio = 1, seed = 100)

# SMOTE will equalize outcome variable

# basic + no lesions + [factor -> lencode_glm] of hospital_number
recipe_base_no_lesions_lecode_hosp_number <-
  recipe(outcome ~ ., data = df_train) %>%
  update_role(id, new_role = "ID") %>%
  step_select(-contains("lesion_")) %>%
  step_mutate(hospital_number = as_factor(hospital_number)) %>% 
  step_lencode_glm(hospital_number, outcome = "outcome") %>% 
  step_nzv(all_nominal_predictors()) %>%
  step_dummy(all_nominal_predictors(), one_hot = T) %>%
  step_nzv(all_predictors()) %>%
  step_YeoJohnson(all_numeric_predictors()) %>%
  step_normalize(all_numeric_predictors())

# same but with SMOTE
recipe_base_no_lesions_lecode_hosp_number_smote <-
  recipe_base_no_lesions_lecode_hosp_number %>% 
  step_smote(outcome, over_ratio = 1, seed = 100)

# basic + no lesions + [factor -> embed] of hospital_number
recipe_base_no_lesions_embed_hosp_number <-
  recipe(outcome ~ ., data = df_train) %>%
  update_role(id, new_role = "ID") %>%
  step_select(-contains("lesion_")) %>%
  step_mutate(hospital_number = as_factor(hospital_number)) %>% 
  step_embed(hospital_number, outcome = "outcome") %>% 
  step_nzv(all_nominal_predictors()) %>%
  step_dummy(all_nominal_predictors(), one_hot = T) %>%
  step_nzv(all_predictors()) %>%
  step_YeoJohnson(all_numeric_predictors()) %>%
  step_normalize(all_numeric_predictors())

# # same but with SMOTE
recipe_base_no_lesions_embed_hosp_number_smote <-
  recipe_base_no_lesions_embed_hosp_number %>% 
  step_smote(outcome, over_ratio = 1, seed = 100)

# basic + embed lesions + embed hospital_number
recipe_base_embed_lesions_hosp_number <-
  recipe(outcome ~ ., data = df_train) %>%
  update_role(id, new_role = "ID") %>%
  step_mutate(across(contains("lesion_"), as.character)) %>% 
  step_mutate(lesions_123 = paste(lesion_1, lesion_2, lesion_3, sep="_")) %>% 
  step_select(-contains("lesion_")) %>%
  step_mutate(hospital_number = as_factor(hospital_number)) %>% 
  step_embed(hospital_number, outcome = "outcome") %>% 
  step_embed(lesions_123, outcome = "outcome") %>% 
  step_nzv(all_nominal_predictors()) %>%
  step_dummy(all_nominal_predictors(), one_hot = T) %>%
  step_nzv(all_predictors()) %>%
  step_YeoJohnson(all_numeric_predictors()) %>%
  step_normalize(all_numeric_predictors())

recipe_base_lencode_lesions_embed_hosp_number <-
  recipe(outcome ~ ., data = df_train) %>%
  update_role(id, new_role = "ID") %>%
  step_mutate(across(contains("lesion_"), as.factor)) %>% 
  step_mutate(lesions_123 = paste(lesion_1, lesion_2, lesion_3, sep="_")) %>% 
  step_select(-contains("lesion_")) %>%
  step_mutate(hospital_number = as_factor(hospital_number)) %>% 
  step_embed(hospital_number, outcome = "outcome") %>% 
  step_lencode_glm(lesions_123, outcome = "outcome") %>% 
  step_nzv(all_nominal_predictors()) %>%
  step_dummy(all_nominal_predictors(), one_hot = T) %>%
  step_nzv(all_predictors()) %>%
  step_YeoJohnson(all_numeric_predictors()) %>%
  step_normalize(all_numeric_predictors())

recipe_base_pooling <- recipe(outcome ~ ., data = df_train) %>%
  update_role(id, new_role = "ID") %>%
  step_other(all_nominal_predictors()) %>% 
  #step_nzv(all_nominal_predictors()) %>% 
  step_dummy(all_nominal_predictors(), one_hot = T) %>%
  step_nzv(all_predictors()) %>%
  step_YeoJohnson(all_numeric_predictors()) %>%
  step_normalize(all_numeric_predictors())


```


## LR standalone

```{r}
data_synth %>% select(where(is.factor)) %>% 
  summary

# lencode all that has > 4 categories (remember step_other!)
recipe_base_lencode4 <- recipe(outcome ~ ., data = df_train) %>%
  update_role(id, new_role = "ID") %>%
  step_other(all_nominal_predictors()) %>% 
  step_zv(all_predictors()) %>%
  step_lencode_glm(hospital_number, outcome = "outcome") %>% 
  step_lencode_glm(temp_of_extremities, outcome = "outcome") %>% 
  step_lencode_glm(peripheral_pulse, outcome = "outcome") %>% 
  step_lencode_glm(mucous_membrane, outcome = "outcome") %>% 
  step_lencode_glm(pain, outcome = "outcome") %>% 
  step_lencode_glm(peristalsis, outcome = "outcome") %>% 
  step_lencode_glm(abdominal_distention, outcome = "outcome") %>% 
  step_lencode_glm(rectal_exam_feces, outcome = "outcome") %>% 
  step_lencode_glm(abdomen, outcome = "outcome") %>% 
  step_dummy(all_nominal_predictors(), one_hot = T) %>% 
  step_YeoJohnson(all_numeric_predictors()) %>%
  step_normalize(all_numeric_predictors()) 
  

lr_spec <- multinom_reg(penalty = tune(),
                        mixture = tune()) %>% 
  set_engine("glmnet", num.threads=6)

lr_lencode4_wf <- workflow() %>% 
  add_model(lr_spec) %>% 
  add_recipe(recipe_base_lencode4)

lr_lencode4_resamples <- lr_lencode4_wf %>%
  tune_grid(
    grid = 20,
    resamples = cv_folds,
    control = control_grid(save_pred = TRUE,
                           save_workflow = TRUE),
    metrics = metric_set(f1_micro)
  )

saveRDS(lr_lencode4_resamples, "models/lr_lencode4_res_synth_only.rds")

autoplot(lr_lencode4_resamples)
```

```{r}
show_best(lr_lencode4_resamples, metric = "f1_micro")
```


## Workflow sets

### LR

```{r}
lr_spec <- multinom_reg(penalty = tune(),
                        mixture = tune()) %>% 
  set_engine("glmnet", num.threads=6)

model_specs_list <- list(LR = lr_spec)

recipes_list <- list(base = recipe_base,
                     pool = recipe_base_pooling,
                     no = recipe_base_no_lesions_hosp_number, 
                     lencode1 =recipe_base_no_lesions_lecode_hosp_number, 
                     smote1 = recipe_base_no_lesions_lecode_hosp_number_smote, 
                     embed = recipe_base_no_lesions_embed_hosp_number, 
                     smote2 = recipe_base_no_lesions_embed_hosp_number_smote,
                     lencode2 = recipe_base_lencode_lesions_embed_hosp_number)

models_set_lr <-
  workflow_set(preproc = recipes_list,
               models = model_specs_list,
               cross = TRUE)

set.seed(124)

lr_recipes_set <-
  models_set_lr %>%
  workflow_map(
    "tune_grid",
    resamples = cv_folds,
    grid = 20,
    metrics = metric_set(f1_micro),
    verbose = TRUE,
    control = control_grid(save_pred = TRUE,
                           save_workflow = TRUE)
  )
saveRDS(lr_recipes_set, "models/lr_wfset_res.rds")

autoplot(lr_recipes_set, select_best = T)
```

```{r}
rank_results(lr_recipes_set, rank_metric = "f1_micro", select_best = TRUE) %>% 
  select(wflow_id, mean, std_err, rank)
```

So, the best recipe is `recipe_base_no_lesions_hosp_number` and `recipe_base_no_lesions_embed_hosp_number`


## GBM with several recipes

### Base

```{r}
library(bonsai)
library(finetune)

gbt_spec <- boost_tree(
  mtry = tune(),
  trees = tune(),
  tree_depth = tune(),
  learn_rate = tune(),
  min_n = tune(),
  loss_reduction = tune(), 
  sample_size = tune(), 
  stop_iter = tune()
) %>%
  set_engine("lightgbm", num.threads=8) %>%
  set_mode("classification") %>%
  translate()

set.seed(124)

gbt_wf_base <-
  workflow(preprocessor = recipe_base, 
           spec = gbt_spec) 

param_set <- extract_parameter_set_dials(gbt_wf_base) %>%
    finalize(x = df_train %>% select(-outcome))

lgb_base_wl_res <- gbt_wf_base %>%
  tune_race_win_loss(
    param_info = param_set,
    resamples = cv_folds,
    grid = 30,
    metrics = metric_set(f1_micro),
    control = control_race(
      verbose_elim = TRUE,
      save_pred = TRUE,
      save_workflow = TRUE,
      burn_in = 10
    )
  )

saveRDS(lgb_base_wl_res, "models/lgb_base_wl_res_synth_only.rds")

plot_race(lgb_base_wl_res)
```


```{r}
show_best(lgb_base_wl_res, metric="f1_micro")
```


### nlnhn - no lesions and no hospital number

```{r}


# model_specs_list <- list(BT = gbt_spec)
# 
# recipes_list <- list(recipe_base, 
#                      recipe_base_no_lesions_hosp_number, 
#                      recipe_base_no_lesions_lecode_hosp_number, 
#                      recipe_base_no_lesions_embed_hosp_number)
# 
# wf_set_gbt <-
#   workflow_set(preproc = recipes_list,
#                models = model_specs_list,
#                cross = TRUE)

set.seed(124)

# models_set_gbt <- models_set_gbt %>%
#   workflow_map(
#     fn = "tune_race_win_loss",
#     param_info = param_set,
#     resamples = cv_folds,
#     grid = 20,
#     metrics = metric_set(f1_micro),
#     verbose = TRUE,
#     control = control_grid(save_pred = TRUE,
#                            save_workflow = TRUE)
#   )

# nlnhn - no lesions and no hospital number
gbt_wf_nlnhn <-
  workflow(preprocessor = recipe_base_no_lesions_hosp_number, 
           spec = gbt_spec) 

param_set <- extract_parameter_set_dials(gbt_wf_nlnhn) %>%
    finalize(x = df_train %>% select(-outcome))

lgb_nlnhn_wl_res <- gbt_wf_nlnhn %>%
  tune_race_win_loss(
    param_info = param_set,
    resamples = cv_folds,
    grid = 30,
    metrics = metric_set(f1_micro),
    control = control_race(
      verbose_elim = TRUE,
      save_pred = TRUE,
      save_workflow = TRUE,
      burn_in = 10
    )
  )

saveRDS(lgb_nlnhn_wl_res, "models/lgb_nlnhn_wl_res_synth_only.rds")
```


```{r}
plot_race(lgb_nlnhn_wl_res)
```

```{r}
show_best(lgb_nlnhn_wl_res, metric="f1_micro")
```

### nlehn - no_lesions_embed_hosp_number

```{r}
set.seed(124)

# nlehn - no lesions and embed hospital number
gbt_wf_nlehn <-
  workflow(preprocessor = recipe_base_no_lesions_embed_hosp_number, 
           spec = gbt_spec) 

param_set <- extract_parameter_set_dials(gbt_wf_nlehn) %>%
    finalize(x = df_train %>% select(-outcome))

lgb_nlehn_wl_res <- gbt_wf_nlehn %>%
  tune_race_win_loss(
    param_info = param_set,
    resamples = cv_folds,
    grid = 30,
    metrics = metric_set(f1_micro),
    control = control_race(
      verbose_elim = TRUE,
      save_pred = TRUE,
      save_workflow = TRUE,
      burn_in = 10
    )
  )

saveRDS(lgb_nlehn_wl_res, "models/lgb_nlehn_wl_res_synth_only.rds")
```


```{r}
plot_race(lgb_nlehn_wl_res)
```

```{r}
show_best(lgb_nlehn_wl_res, metric="f1_micro")
```

### elehn - embed joined lesions and hospital number

```{r}
set.seed(124)

# elehn - embed lesions and embed hospital number
gbt_wf_elehn <-
  workflow(preprocessor = recipe_base_embed_lesions_hosp_number, 
           spec = gbt_spec) 

param_set <- extract_parameter_set_dials(gbt_wf_elehn) %>%
    finalize(x = df_train %>% select(-outcome))

lgb_elehn_wl_res <- gbt_wf_elehn %>%
  tune_race_win_loss(
    param_info = param_set,
    resamples = cv_folds,
    grid = 30,
    metrics = metric_set(f1_micro),
    control = control_race(
      verbose_elim = TRUE,
      save_pred = TRUE,
      save_workflow = TRUE,
      burn_in = 10
    )
  )

saveRDS(lgb_elehn_wl_res, "models/lgb_elehn_wl_res_synth_only.rds")

plot_race(lgb_elehn_wl_res)
```

### llehn - lencode joined lesions and embed hospital number

```{r}
set.seed(124)

# llehn - lencode lesions and embed hospital number
gbt_wf_llehn <-
  workflow(preprocessor = recipe_base_lencode_lesions_embed_hosp_number, 
           spec = gbt_spec) 

param_set <- extract_parameter_set_dials(gbt_wf_llehn) %>%
    finalize(x = df_train %>% select(-outcome))

lgb_llehn_wl_res <- gbt_wf_llehn %>%
  tune_race_win_loss(
    param_info = param_set,
    resamples = cv_folds,
    grid = 30,
    metrics = metric_set(f1_micro),
    control = control_race(
      verbose_elim = TRUE,
      save_pred = TRUE,
      save_workflow = TRUE,
      burn_in = 10
    )
  )

saveRDS(lgb_llehn_wl_res, "models/lgb_llehn_wl_res_synth_only.rds")

plot_race(lgb_llehn_wl_res)
```
```{r}
show_best(lgb_llehn_wl_res, metric="f1_micro")
```

### bp - Base and pooling of rare categories (step_other)

```{r}
set.seed(124)

# bp - base + pooling
gbt_wf_bp <-
  workflow(preprocessor = recipe_base_pooling, 
           spec = gbt_spec) 

param_set <- extract_parameter_set_dials(gbt_wf_bp) %>%
    finalize(x = df_train %>% select(-outcome))

lgb_bp_wl_res <- gbt_wf_bp %>%
  tune_race_win_loss(
    param_info = param_set,
    resamples = cv_folds,
    grid = 30,
    metrics = metric_set(f1_micro),
    control = control_race(
      verbose_elim = TRUE,
      save_pred = TRUE,
      save_workflow = TRUE,
      burn_in = 10
    )
  )

saveRDS(lgb_bp_wl_res, "models/lgb_bp_wl_res_synth_only.rds")

plot_race(lgb_bp_wl_res)
```

```{r}
show_best(lgb_bp_wl_res, metric="f1_micro")
```

### bnd - base but no dummies

```{r}
recipe_base_nodum <- recipe(outcome ~ ., data = df_train) %>%
  update_role(id, new_role = "ID") %>%
  step_other(all_nominal_predictors()) %>% 
  step_zv(all_predictors()) %>%
  step_YeoJohnson(all_numeric_predictors()) %>%
  step_normalize(all_numeric_predictors()) 

set.seed(124)

gbt_wf_nodum <-
  workflow(preprocessor = recipe_base_nodum, 
           spec = gbt_spec) 

param_set <- extract_parameter_set_dials(gbt_wf_nodum) %>%
    finalize(x = df_train %>% select(-outcome))

lgb_nodum_wl_res <- gbt_wf_nodum %>%
  tune_race_win_loss(
    param_info = param_set,
    resamples = cv_folds,
    grid = 30,
    metrics = metric_set(f1_micro),
    control = control_race(
      verbose_elim = TRUE,
      save_pred = TRUE,
      save_workflow = TRUE,
      burn_in = 10
    )
  )

saveRDS(lgb_nodum_wl_res, "models/lgb_nodum_wl_res_synth_only.rds")

plot_race(lgb_nodum_wl_res)
```

```{r}
show_best(lgb_nodum_wl_res)
```

### bpnd - base and pooling and no dummies, lencode hospital

```{r}
recipe_base_nodum_pool <- recipe(outcome ~ ., data = df_train) %>%
  update_role(id, new_role = "ID") %>%
  step_mutate(across(contains("lesion_"), as.factor)) %>%
  step_mutate(hospital_number = as_factor(hospital_number)) %>%
  step_lencode_glm(hospital_number, outcome = "outcome") %>%
  step_other(all_nominal_predictors()) %>%
  step_nzv(all_predictors()) %>%
  step_normalize(all_numeric_predictors())

gbt_spec <- boost_tree(
  mtry = tune(),
  trees = tune(),
  tree_depth = tune(),
  learn_rate = tune(),
  min_n = tune(),
  loss_reduction = tune()
) %>%
  set_engine("lightgbm", num.threads=8) %>%
  set_mode("classification") %>%
  translate()

set.seed(124)

gbt_wf_nodum_pool <-
  workflow(preprocessor = recipe_base_nodum_pool, 
           spec = gbt_spec) 

param_set <- extract_parameter_set_dials(gbt_wf_nodum_pool) %>%
    finalize(x = df_train %>% select(-outcome))

lgb_nodum_pool_wl_res <- gbt_wf_nodum_pool %>%
  tune_race_win_loss(
    param_info = param_set,
    resamples = cv_folds,
    grid = 30,
    metrics = metric_set(f1_micro),
    control = control_race(
      verbose_elim = TRUE,
      save_pred = TRUE,
      save_workflow = TRUE,
      burn_in = 10
    )
  )

saveRDS(lgb_nodum_pool_wl_res, "models/lgb_nodum_pool_wl_res_synth_only.rds")

plot_race(lgb_nodum_pool_wl_res)
```

### From a notebook

> df['lesion_2'] = df['lesion_2'].apply(lambda x:1 if x>0 else 0)

> df["deviation_from_normal_temp"] = df["rectal_temp"].apply(lambda x: abs(x - 37.8))


```{r}
recipe_base_notebook <- recipe(outcome ~ ., data = df_train) %>%
  update_role(id, new_role = "ID") %>%
  step_select(-lesion_3) %>% 
  step_mutate(lesion_2 = if_else(lesion_2 > 0, 1, 0)) %>% 
  step_mutate(hospital_number = as_factor(hospital_number)) %>%
  step_lencode_glm(hospital_number, outcome = "outcome") %>%
  step_mutate(deviation_from_normal_temp = abs(rectal_temp - 37.8)) %>% 
  step_novel(all_nominal_predictors()) %>% 
  step_other(all_nominal_predictors()) %>% 
  step_zv(all_predictors()) %>%
  step_YeoJohnson(all_numeric_predictors()) %>%
  step_normalize(all_numeric_predictors()) 

gbt_spec <- boost_tree(
  mtry = tune(),
  trees = tune(),
  tree_depth = tune(),
  learn_rate = tune(),
  min_n = tune(),
  loss_reduction = tune()
) %>%
  set_engine("lightgbm", num.threads=8) %>%
  set_mode("classification") %>%
  translate()

set.seed(124)

gbt_wf_notebook <-
  workflow(preprocessor = recipe_base_notebook, 
           spec = gbt_spec) 

param_set <- extract_parameter_set_dials(gbt_wf_notebook) %>%
    finalize(x = df_train %>% select(-outcome))

lgb_wl_notebook_res <- gbt_wf_notebook %>%
  tune_race_win_loss(
    param_info = param_set,
    resamples = cv_folds,
    grid = 30,
    metrics = metric_set(f1_micro),
    control = control_race(
      verbose_elim = TRUE,
      save_pred = TRUE,
      save_workflow = TRUE,
      burn_in = 10
    )
  )

plot_race(lgb_wl_notebook_res)
```



# Modelling on joined data

I'll keep NAs as is because GBT can handle them

```{r}
data_full <- bind_rows(data_synth %>% select(-id), data_orig) %>% 
  filter(complete.cases(.))

data_full %>% 
  skim()
```
## Split (non-stratified)

```{r}
set.seed(124)

data_split <- initial_split(data_full, prop = 0.99)

df_train <- training(data_split)
df_test <- testing(data_split)

cv_folds <- vfold_cv(df_train, v = 10, repeats = 5) 

f1_micro <- metric_tweak("f1_micro", f_meas, estimator = "micro")
```

## Recipes

Two best recipes: one for LE, one for GBT

LR: `recipe_base_lencode_lesions_embed_hosp_number`

GBT: `recipe_base_nodum``

```{r}
# for LR
recipe_base_lencode_lesions_embed_hosp_number <-
  recipe(outcome ~ ., data = df_train) %>%
  step_mutate(across(contains("lesion_"), as.factor)) %>% 
  step_mutate(lesions_123 = paste(lesion_1, lesion_2, lesion_3, sep="_")) %>% 
  step_select(-contains("lesion_")) %>%
  step_mutate(hospital_number = as_factor(hospital_number)) %>% 
  step_embed(hospital_number, outcome = "outcome") %>% 
  step_lencode_glm(lesions_123, outcome = "outcome") %>% 
  step_nzv(all_nominal_predictors()) %>%
  step_dummy(all_nominal_predictors(), one_hot = T) %>%
  step_nzv(all_predictors()) %>%
  step_YeoJohnson(all_numeric_predictors()) %>%
  step_normalize(all_numeric_predictors())

# I'll add step_novel() here
recipe_base_nodum <- recipe(outcome ~ ., data = df_train) %>%
  step_novel(all_nominal_predictors()) %>% 
  step_other(all_nominal_predictors()) %>% 
  step_zv(all_predictors()) %>%
  step_YeoJohnson(all_numeric_predictors()) %>%
  step_normalize(all_numeric_predictors()) 
```

## Multinomial LR

```{r}
lr_spec <- multinom_reg(penalty = tune(),
                        mixture = tune()) %>% 
  set_engine("glmnet", num.threads=6)

lr_wf <- workflow() %>% 
  add_model(lr_spec) %>% 
  add_recipe(recipe_base_lencode_lesions_embed_hosp_number)

lr_full_res <- lr_wf %>%
  tune_grid(
    grid = 20,
    resamples = cv_folds,
    control = control_grid(save_pred = TRUE,
                           save_workflow = TRUE),
    metrics = metric_set(f1_micro)
  )

saveRDS(lr_full_res, "models/lr_bllehn_full_data.rds")

autoplot(lr_full_res)
```

```{r}
show_best(lr_full_res, metric = "f1_micro")
```

## GBT

```{r}
library(finetune)
library(bonsai)

gbt_spec <- boost_tree(
  mtry = tune(),
  trees = tune(),
  tree_depth = tune(),
  learn_rate = tune(),
  min_n = tune(),
  loss_reduction = tune()
) %>%
  set_engine("lightgbm", num.threads=8) %>%
  set_mode("classification") %>%
  translate()

set.seed(124)

gbt_wf <-
  workflow(preprocessor = recipe_base_nodum, 
           spec = gbt_spec) 

param_set <- extract_parameter_set_dials(gbt_wf) %>%
    finalize(x = df_train %>% select(-outcome))

gbt_full_res_race <- gbt_wf %>%
  tune_race_win_loss(
    param_info = param_set,
    resamples = cv_folds,
    grid = 20,
    metrics = metric_set(f1_micro, mcc),
    control = control_race(
      verbose_elim = TRUE,
      save_pred = TRUE,
      save_workflow = TRUE,
      burn_in = 10
    )
  )

saveRDS(gbt_full_res_race, "models/gbt_race_full_data.rds")

plot_race(gbt_full_res_race)
```

```{r}
show_best(gbt_full_res_race)
```

### Bayes

```{r}
gbt_full_bres <-
      tune_bayes(
        gbt_wf,
        resamples = cv_folds,
        initial = 24,
        iter = 30,
        metrics = metric_set(f1_micro),
        param_info = param_set,
        control = control_bayes(
          no_improve = 20,
          save_pred = TRUE,
          verbose = FALSE,
          save_workflow = TRUE
        )
      )

show_best(gbt_full_bres)
```


## Prediction

```{r}
data_test <- read_csv("data/test.csv", show_col_types = F) %>% 
  mutate(across(is_character, ~ as_factor(.x)),
         hospital_number = as_factor(hospital_number))

data_test
```

### GBT

```{r}
gbt_fit <- fit_best(gbt_full_res_race)
gbt_fit <- fit_best(gbt_full_bres)

data_test_proc <- recipe_base_nodum %>% 
  prep() %>% 
  bake(new_data = data_test)

# it seems, here predict does the processing, otherwise I don't know how to explain why
# public score on data_test_proc is 0.60 vs 0.80 on data_set
test_prediction <- predict(gbt_fit, data_test) %>% 
  mutate(id = data_test$id, .before = .pred_class) %>% 
  rename("outcome" = .pred_class)

write_csv(x =  test_prediction, file = "data/test_prediction_bayes.csv")
```

