---
title: "smokers"
author: "AG"
format: html
editor: source
---

## Data

```{r, message=FALSE}
library(tidyverse)
library(tidymodels)
library(ggpubr)
library(finetune)
library(bonsai)
library(embed)
library(themis)
library(baguette)

plot_validation_results <- function(dat, components) {
  # dat - data table
  # components - vector of components names to plot
  dat %>%
    select(-id) %>%
    select(all_of(c(components, "smoking"))) %>% 
    # Create the scatterplot matrix
    ggplot(aes(x = .panel_x, y = .panel_y, color = smoking, fill = smoking)) +
    geom_point(alpha = 0.2, size = 1) +
    ggforce::geom_autodensity(alpha = .3) +
    ggforce::facet_matrix(vars(-smoking), layer.diag = 2) + 
    scale_color_brewer(palette = "Dark2") + 
    scale_fill_brewer(palette = "Dark2")
}

data_train <- read_csv("data/train.csv.zip") %>% 
  mutate(smoking = as_factor(if_else(smoking == 0, "no", "yes")))

data_real <- read_csv("data/train_dataset.csv.zip") %>% 
  mutate(smoking = as_factor(if_else(smoking == 0, "no", "yes")))

data_full <- bind_rows(data_train, data_real)

setdiff(names(data_train), names(data_real))
```

```{r}
data_train %>% skimr::skim()
```

## EDA

### Histograms as is

```{r, fig.height=10}
plot_hist <- function(varname, df){
  df %>% 
    ggplot(aes(.data[[varname]])) +
    geom_histogram(aes(fill=smoking), bins = 100) +
    geom_rug()
}

hist_list <- map(names(data_train)[2:23], ~ plot_hist(., data_train))

ggarrange(plotlist = hist_list, nrow = 8, ncol = 3, common.legend = T)
```

Let's apply some transformations

### Histograms after transform

```{r, fig.height=10}
recipe_log_trans <- recipe(smoking ~ ., data = data_train) %>% 
  update_role(id, new_role = "ID") %>% 
  step_zv(all_predictors()) %>% 
  step_log(all_numeric_predictors(), offset = 1) %>% 
  step_normalize(all_numeric_predictors())

data_log <- recipe_log_trans %>% prep() %>% juice()

hist_list <- map(names(data_log)[2:23], ~ plot_hist(., data_log))

ggarrange(plotlist = hist_list, nrow = 8, ncol = 3, common.legend = T)
```

### Histograms after YJ transform

```{r, fig.height=10}
recipe_yj_trans <- recipe(smoking ~ ., data = data_train) %>% 
  update_role(id, new_role = "ID") %>% 
  step_zv(all_predictors()) %>% 
  step_YeoJohnson(all_numeric_predictors()) %>% 
  step_normalize(all_numeric_predictors())

data_yj <- recipe_yj_trans %>% prep() %>% juice()

hist_list <- map(names(data_yj)[2:23], ~ plot_hist(., data_yj))

ggarrange(plotlist = hist_list, nrow = 8, ncol = 3, common.legend = T)
```

### Boxplots

```{r, fig.height=18}
plot_box <- function(varname, df){
  df %>% 
    ggplot(aes(smoking, .data[[varname]])) +
    geom_boxplot(aes(fill=smoking), notch = F) +
    xlab("")
}

box_list <- map(names(data_log)[2:23], ~ plot_box(., data_log))

ggarrange(plotlist = box_list, nrow = 6, ncol = 4, common.legend = T)

```

### Corrplot

```{r, fig.width=8}
library(ggcorrplot)

corr_matrix <- round(cor(data_train[,-c(1, 24)]), 1) 

ggcorrplot(corr_matrix,
           hc.order = TRUE,
           type = "upper",
           lab = TRUE)
```

```{r, fig.width=9}
corr_matrix <- round(cor(data_yj[,-c(1, 24)]), 1) 

ggcorrplot(corr_matrix,
           hc.order = TRUE,
           type = "upper",
           lab = TRUE)
```

### UMAP

In the recipes below I do down-sampling

#### Non-supervised

```{r, fig.height=8, fig.width=8}
umap_yj_recipe <- 
  recipe(smoking ~ ., data = data_train) %>%
  update_role(id, new_role = "ID") %>%
  step_nzv(all_predictors()) %>% 
  step_YeoJohnson(all_numeric_predictors()) %>% 
  step_normalize(all_numeric_predictors()) %>%
  step_downsample(smoking, under_ratio = 1, seed = 99) %>% 
  step_umap(
    all_numeric_predictors(),
    num_comp = 10,
    min_dist = 0.8,
    neighbors = 25
  ) %>% 
  step_normalize(all_numeric_predictors())

umap_results <- umap_yj_recipe %>%
  prep() %>% 
  juice() 


umap_results %>% 
  slice_sample(n = 1000, by = smoking) %>% 
  plot_validation_results(c("UMAP01", "UMAP02", "UMAP03", "UMAP04")) +
  ggtitle("UMAP unsupervised")
```

#### Supervised

```{r, fig.height=8, fig.width=8}
umap_sup_recipe <-
  recipe(smoking ~ ., data = data_train) %>%
  update_role(id, new_role = "ID") %>%
  step_nzv(all_predictors()) %>%
  step_YeoJohnson(all_numeric_predictors()) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_downsample(smoking, under_ratio = 1, seed = 99) %>% 
  step_umap(
    all_numeric_predictors(),
    outcome = "smoking",
    num_comp = 10,
    min_dist = 0.8,
    neighbors = 25
  )

umap_sup_results <- umap_sup_recipe %>%
  prep() %>% 
  juice() 


umap_sup_results %>% 
  slice_sample(n = 1000, by = smoking) %>% 
  plot_validation_results(c("UMAP01", "UMAP02", "UMAP03", "UMAP04")) +
  ggtitle("UMAP supervised")
```

### PCA

```{r, fig.height=8, fig.width=8}
rec_pca <- 
  recipe(smoking ~ ., data = data_train) %>%
  update_role(id, new_role = "ID") %>%
  step_nzv(all_predictors()) %>%
  step_YeoJohnson(all_numeric_predictors()) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_downsample(smoking, under_ratio = 1, seed = 99) %>% 
  step_pca(all_numeric_predictors(), num_comp = 8) %>% 
  step_normalize(all_numeric_predictors())

rec_pca %>% 
  prep() %>% 
  juice() %>% 
  slice_sample(n = 1000, by = smoking) %>% 
  plot_validation_results(c("PC1", "PC2", "PC3", "PC4", "PC5")) +
  ggtitle("Principal Component Analysis")
```

## Data split, CV folds, metric, basic recipe

```{r}
library(themis)

# split
set.seed(124)

data_split <- initial_split(data_full, prop = 0.99, strata = smoking)

df_train <- training(data_split)
df_test <- testing(data_split)

# folds
cv_folds <- vfold_cv(df_train, strata = "smoking", v = 10, repeats = 5) 

metrics <- metric_set(roc_auc)

# recipe
rec_base <-
  recipe(smoking ~ ., data = df_train) %>%
  update_role(id, new_role = "ID") %>%
  step_nzv(all_predictors()) %>%
  step_downsample(smoking, under_ratio = 1, seed = 100) %>% 
  step_YeoJohnson(all_numeric_predictors()) %>%
  step_normalize(all_numeric_predictors()) 

rec_base %>% prep() %>% juice() %>% group_by(smoking) %>% count()
```

## LR basic

```{r}
lr_spec <- logistic_reg(penalty = tune(),
                        mixture = tune()) %>%
  set_engine("glmnet")

lr_wf <- workflow(preprocessor = rec_base,
                  spec = lr_spec)

lr_res <- lr_wf %>% 
  tune_grid(
    grid = 20,
    resamples = cv_folds,
    control = control_grid(save_pred = TRUE,
                           save_workflow = FALSE),
    metrics = metric_set(roc_auc)
  )

lr_res %>% show_best()

lr_best_param <- select_best(lr_res)
lr_model_final <- finalize_model(lr_spec, lr_best_param)
# should data set be processed here?
# proc and not-proc fit objects are not identical
lr_fit_proc <- fit(lr_model_final, smoking ~ ., rec_base %>% prep() %>% juice())

saveRDS(lr_fit_proc, "models/lr_base_fit_proc.rds")

lr_res %>% show_best()
```

```{r}
data_test <- read_csv("data/test.csv.zip", show_col_types = F)

data_test_proc <- rec_base %>% prep() %>% bake(new_data = data_test)

test_prediction <- predict(lr_fit_proc, data_test_proc, type="prob") %>% 
  mutate(id = data_test$id, .before = .pred_yes) %>% 
  rename("smoking" = .pred_yes) %>% 
  select(-.pred_no)

write_csv(x =  test_prediction, file = "submissions/test_prediction_lr_base.csv")
```

Public AUC = 0.84235

## GBT basic + racing

```{r}
gbt_spec <- boost_tree(
  trees = tune(),
  tree_depth = tune(),
  learn_rate = tune(),
  min_n = tune(),
  loss_reduction = tune(), 
  sample_size = tune(), 
  stop_iter = tune()
) %>%
  set_engine("lightgbm", num.threads=8) %>%
  set_mode("classification") %>%
  translate()

set.seed(124)

gbt_wf_base <-
  workflow(preprocessor = rec_base, 
           spec = gbt_spec) 

param_set <- extract_parameter_set_dials(gbt_wf_base) %>%
    finalize(x = df_train %>% select(-smoking))

gbt_base_wl_res <- gbt_wf_base %>%
  tune_race_win_loss(
    param_info = param_set,
    resamples = cv_folds,
    grid = 40,
    metrics = metric_set(roc_auc),
    control = control_race(
      verbose_elim = TRUE,
      save_pred = TRUE,
      save_workflow = FALSE,
      burn_in = 10
    )
  )

show_best(gbt_base_wl_res)
```

### Prediction

```{r}
gbt_best_param <- select_best(gbt_base_wl_res)
gbt_model_final <- finalize_model(gbt_spec, gbt_best_param)
# should data set be processed here?
# proc and not-proc fit objects are not identical
gbt_fit_proc <- fit(gbt_model_final, smoking ~ ., rec_base %>% prep() %>% juice())

saveRDS(gbt_fit_proc, "models/gbt_base_fit_proc.rds")

test_prediction <- predict(gbt_fit_proc, data_test_proc, type="prob") %>% 
  mutate(id = data_test$id, .before = .pred_yes) %>% 
  rename("smoking" = .pred_yes) %>% 
  select(-.pred_no)

write_csv(x =  test_prediction, file = "submissions/test_prediction_gbt_base.csv")
```

Public AUC = 0.86828

## GBT with grid search and Bayes opt

```{r}
gbt_spec <- boost_tree(
  trees = tune(),
  tree_depth = tune(),
  learn_rate = tune(),
  min_n = tune(),
  loss_reduction = tune(), 
  sample_size = tune(), 
  stop_iter = tune()
) %>%
  set_engine("lightgbm", num.threads=8) %>%
  set_mode("classification") %>%
  translate()

set.seed(124)

gbt_wf_base <-
  workflow(preprocessor = rec_base, 
           spec = gbt_spec) 

param_set <- extract_parameter_set_dials(gbt_wf_base) %>%
    finalize(x = df_train %>% select(-smoking))

gbt_base_res <- gbt_wf_base %>%
  tune_grid(
    grid = 40,
    param_info = param_set,
    resamples = cv_folds,
    control = control_grid(save_pred = TRUE,
                           save_workflow = FALSE),
    metrics = metric_set(roc_auc)
  )



# Bayes
gbt_base_bres <-
      tune_bayes(
        gbt_wf_base,
        resamples = cv_folds,
        initial = gbt_base_res,
        iter = 30,
        metrics = metric_set(roc_auc),
        param_info = param_set,
        control = control_bayes(
          no_improve = 20,
          save_pred = TRUE,
          verbose = FALSE,
          save_workflow = FALSE
        )
      )


show_best(gbt_base_res)
```

```{r}
show_best(gbt_base_bres)
```

### Prediction

```{r}
data_test <- read_csv("data/test.csv.zip", show_col_types = F)
data_test_proc <- rec_base %>% prep() %>% bake(new_data = data_test)

gbt_best_param <- select_best(gbt_base_bres)
gbt_model_final <- finalize_model(gbt_spec, gbt_best_param)
# should data set be processed here?
# proc and not-proc fit objects are not identical
gbt_fit_proc <- fit(gbt_model_final, smoking ~ ., rec_base %>% prep() %>% juice())

saveRDS(gbt_fit_proc, "models/gbt_full_data_baseB_fit_proc.rds")

test_prediction <- predict(gbt_fit_proc, data_test_proc, type="prob") %>% 
  mutate(id = data_test$id, .before = .pred_yes) %>% 
  rename("smoking" = .pred_yes) %>% 
  select(-.pred_no)

write_csv(x =  test_prediction, file = "submissions/test_prediction_gbt_baseB_full_data.csv")
```

Public ROC AUC = 0.87050 (place 360)
Full data set public score = 0.86

## MLP

BAGging takes too long

```{r}
# rec_nocorr <- 
#   recipe(defects ~ ., data = df_train) %>%
#   update_role(id, new_role = "ID") %>%
#   step_nzv(all_predictors()) %>%
#   step_log(all_numeric_predictors(), offset = 1) %>%
#   step_normalize(all_numeric_predictors()) %>% 
#   step_downsample(defects, under_ratio = 1, seed = 100) %>% 
#   step_corr(all_predictors(), threshold = tune("corr_tune"))

mlp_spec <- mlp(
      hidden_units = tune(),
      penalty = tune(),
      epochs = tune()) %>%
      set_engine("nnet", num.threads = 4) %>%
      set_mode("classification") 

mlp_wf <-
  workflow(preprocessor = rec_base, 
           spec = mlp_spec)

mlp_res <- mlp_wf %>% 
  tune_grid(
    grid = 40,
    resamples = cv_folds,
    control = control_grid(save_pred = TRUE,
                           save_workflow = FALSE),
    metrics = metric_set(roc_auc)
  )

mlp_res %>% show_best()
```

```{r}
data_test <- read_csv("data/test.csv.zip", show_col_types = F)
data_test_proc <- rec_base %>% prep() %>% bake(new_data = data_test)

mlp_best_param <- select_best(mlp_res)
mlp_model_final <- finalize_model(mlp_spec, mlp_best_param)
# should data set be processed here?
# proc and not-proc fit objects are not identical
mlp_fit_proc <- fit(mlp_model_final, smoking ~ ., rec_base %>% prep() %>% juice())

saveRDS(mlp_fit_proc, "models/mlp_base_fit_proc.rds")

test_prediction <- predict(mlp_fit_proc, data_test_proc, type="prob") %>% 
  mutate(id = data_test$id, .before = .pred_yes) %>% 
  rename("smoking" = .pred_yes) %>% 
  select(-.pred_no)

write_csv(x =  test_prediction, file = "submissions/test_prediction_mlp_base.csv")
```

Public score 0.766

## Stacks

### Specs

```{r}
library(stacks)

ctrl_grid <- control_stack_grid()
ctrl_res <- control_stack_resamples()

# model definitions
boost_tree_lgboost_spec <-
  boost_tree(
    tree_depth = tune(),
    trees = tune(),
    learn_rate = tune(),
    min_n = tune(),
    loss_reduction = tune(),
    sample_size = tune(),
    stop_iter = tune()
  ) %>%
  set_engine('lightgbm', num.threads=8) %>%
  set_mode('classification') %>%
  translate()

logistic_reg_glmnet_spec <-
  logistic_reg(penalty = tune(), mixture = tune()) %>%
  set_engine('glmnet', num.threads=8)

mlp_nnet_spec <-
  mlp(hidden_units = tune(), penalty = tune(), epochs = tune()) %>%
  set_engine('nnet', num.threads=8) %>%
  set_mode('classification')
```


### Recipes

```{r, fig.height=6}
# recipes
rec_base <-
  recipe(defects ~ ., data = df_train) %>%
  update_role(id, new_role = "ID") %>%
  step_nzv(all_predictors()) %>%
  step_log(all_numeric_predictors(), offset = 1) %>%
  step_normalize(all_numeric_predictors()) %>% 
  step_downsample(defects, under_ratio = 1, seed = 100)

rec_corr_tune <- 
  recipe(defects ~ ., data = df_train) %>%
  update_role(id, new_role = "ID") %>%
  step_nzv(all_predictors()) %>%
  step_log(all_numeric_predictors(), offset = 1) %>%
  step_normalize(all_numeric_predictors()) %>% 
  step_downsample(defects, under_ratio = 1, seed = 100) %>% 
  step_corr(all_predictors(), threshold = tune("corr_tune"))

rec_corr <-
  recipe(defects ~ ., data = df_train) %>%
  update_role(id, new_role = "ID") %>%
  step_nzv(all_predictors()) %>%
  step_log(all_numeric_predictors(), offset = 1) %>%
  step_normalize(all_numeric_predictors()) %>% 
  step_downsample(defects, under_ratio = 1, seed = 100) %>% 
  step_corr(all_predictors(), threshold = 0.9)

df_train_corr <- rec_corr %>% prep() %>% juice() 

corr_matrix <- round(cor(df_train_corr[,-c(1, ncol(df_train_corr))]), 1) 

# corr threshold 0.8 leaves only 6 predictors!
ggcorrplot(corr_matrix,
           hc.order = TRUE,
           type = "upper",
           lab = TRUE)
```

### Workflows

```{r}
lr_wf <- workflow(rec_corr, logistic_reg_glmnet_spec)

mlp_wf <- workflow(rec_corr, mlp_nnet_spec)

gbt_wf <- workflow(rec_corr, boost_tree_lgboost_spec)

```


### Tuning & validation

Via repeated CV, grid of size 20

```{r}

# param_set <- extract_parameter_set_dials(gbt_wf) %>%
#     finalize(x = df_train %>% select(-defects))

gbt_res <- gbt_wf %>%
  tune_grid(
    grid = 20,
    #param_info = param_set,
    resamples = cv_folds,
    control = ctrl_grid,
    metrics = metric_set(roc_auc)
  )

lr_res <- lr_wf %>% 
  tune_grid(
    grid = 20,
    resamples = cv_folds,
    control = ctrl_grid,
    metrics = metric_set(roc_auc)
  )

mlp_res <- mlp_wf %>%
  tune_grid(
    grid = 20,
    resamples = cv_folds,
    control = ctrl_grid,
    metrics = metric_set(roc_auc)
  )


saveRDS(lr_res, "models/lr_res_stacks.rds")
saveRDS(mlp_res, "models/mlp_res_stacks.rds")
saveRDS(gbt_res, "models/gbt_res_stacks.rds")

#lr_res <- readRDS("models/lr_res_stacks.rds")
#mlp_res <- readRDS("models/mlp_res_stacks.rds")
# gbt_res <- readRDS("models/gbt_res_stacks.rds")
```

```{r}
gbt_res %>% show_best()
```

```{r}
mlp_res %>% show_best()
```


### Stacking

```{r}
model_stacks <- stacks() %>% 
  add_candidates(lr_res) %>% 
  add_candidates(mlp_res) %>% 
  add_candidates(gbt_res)

as_tibble(model_stacks)
```

```{r}
model_stack_blend <- model_stacks %>% 
  blend_predictions(penalty = 10^(-6:-1),
                    mixture = 1)

autoplot(model_stack_blend)
```

```{r}
autoplot(model_stack_blend, type = "members")
```

```{r}
autoplot(model_stack_blend, type="weights")
```

### Look at params

```{r}
model_stack_fit <- model_stack_blend %>% fit_members()

collect_parameters(model_stack_fit, "gbt_res")
```

### Prediction

```{r}
data_test <- read_csv("data/test.csv.zip", show_col_types = F)
data_test_proc <- rec_base %>% prep() %>% bake(new_data = data_test)

mlp_best_param <- select_best(mlp_res)
mlp_model_final <- finalize_model(mlp_spec, mlp_best_param)
# should data set be processed here?
# proc and not-proc fit objects are not identical
mlp_fit_proc <- fit(mlp_model_final, smoking ~ ., rec_base %>% prep() %>% juice())

saveRDS(mlp_fit_proc, "models/mlp_base_fit_proc.rds")

test_prediction <- predict(mlp_fit_proc, data_test_proc, type="prob") %>% 
  mutate(id = data_test$id, .before = .pred_yes) %>% 
  rename("smoking" = .pred_yes) %>% 
  select(-.pred_no)

write_csv(x =  test_prediction, file = "submissions/test_prediction_mlp_base.csv")
```


```{r}
data_test <- read_csv("data/test.csv.gz", show_col_types = F)

data_test_proc <- rec_corr %>% prep() %>% bake(new_data = data_test)

# Don't process the test data here
test_prediction <- predict(model_stack_fit, data_test, type="prob") %>% 
  mutate(id = data_test$id, .before = .pred_yes) %>% 
  rename("smoking" = .pred_yes) %>% 
  select(-.pred_no)

#write_csv(x =  test_prediction, file = "submissions/test_prediction_stacks3_corr.csv")
```

