---
title: "Student Performance"
author: "A.G."
date-modified: "last update: `r format(Sys.Date(), format = '%d %B %Y')`"
format: 
  html:
    toc: true
    toc-depth: 2
    toc-title: Contents
    toc-location: left
    df-print: paged
    standalone: true
    code-fold: true
editor: source
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F, cache = F)
library(skimr)
library(tidyverse)
library(caret)
library(DMwR) # installed via archive: install.packages("/Path/to/DMwR_0.4.1.tar.gz", repos=NULL, type="source" )
library(ROCR)
```


# Intro

This data set consists of the marks secured by the students in various subjects.

Example Research Questions:

  - How effective is the test preparation course?
  
  - Which major factors contribute to test outcomes?
  
  - What would be the best way to improve student scores on each test?
  
  - What patterns and interactions in the data can you find? Let me know in the comments section below.

# Read the data

```{r}
df <- read_csv("data/exams.csv", 
               col_types = "fffffddd", 
               name_repair = "universal")

# for better readability remove 'group' in 'race/ethnicity'
df$race.ethnicity <- as.factor(str_replace(df$race.ethnicity, "group ", ""))

df
```

# Descriptive statistics

```{r}
skim_desc <- skim(df)

skim_desc %>%
  yank("factor")
```

```{r}
skim_desc <- skim(df)

skim_desc %>%
  yank("numeric")
```

